---
title: "drugbank_crawl"
author: "姜玟昊"
date: "2022-09-26"
output: html_document
---
last version: 20230602

```{r setup, include=FALSE}
# knitr::opts_chunk$set(echo = TRUE)
# pacman::p_unload(pacman::p_loaded(), character.only = TRUE)
# rm(list = ls())
# gc()
library(tidyverse)
library(magrittr)
library(reticulate)
# setwd('//172.16.13.136/share/members/jiangwenhao/TPHP/20220908/TPHP_carcinoma_upregulated_overlapped_drugbank/')
# setwd('N:/code/6_4drug')
setwd('X:/code/6_4drug')
# source('../source/source_code.R')
# source_python('drugbank_crawl_source.py') # source from python

```


# 1. Clinical trials of all DEPs
## 1.1 筛选targets中包含dysregulated proteins的drugs
```{r}
# drugbank table
# df_drug <- rio::import('//172.16.13.136/share/members/jiangwenhao/TPHP/20220908/drug/drugbank_results_all_20230207.csv')
df_drug <- rio::import('input/drugbank_results_all_20230207.csv')
df_drug1 <- df_drug %>% filter(Biotech == T) # biotech only
df_drug2 <- df_drug1 %>%
  select(`DrugBank ID`:Description, `Target uniprot`) %>%
  mutate(`Target uniprot` = str_replace_all(`Target uniprot`, ' ', '')) %>%
  separate_rows(`Target uniprot`, sep = ';') %>%
  filter(`Target uniprot` != '')
  

# DEA results
df_dea <- rio::import('../6_3paired_tumor_nontumor/output/tumor_nontumor_analysis_reports_v1202.xlsx', sheet = 2)
meta <- rio::import('../6_3paired_tumor_nontumor/output/T_NT_info_check.xlsx')
df_dea1 <- df_dea %>%
  inner_join(meta %>% distinct(cancer, cancer_abbr), .) %>% 
  filter(abs(g_adj) >= 0.5, p_adj_BH < 0.05)

df_map <- df_dea1 %>%
  inner_join(df_drug2, by = c('protein' = 'Target uniprot')) %>% 
  arrange(cancer, Name)


```

## 1.2 crawl clinical trials -- all/with results
```{r}
df_cancer_drug <- df_map %>% select(cancer, Name) %>% distinct()
df_cancer_drug %>% rio::export('output/TPHP_cancer_drug_map_20251202.xlsx')

#检查特殊字符
df_cancer_drug$Name %>% unique %>% str_c(collapse = '') %>% str_replace_all('\\w', '') %>% str_split('') %>% .[[1]] %>% unique() %>% str_c(collapse = '') %>% writeClipboard() # -(,[]) .'
```

### python crawler
```{prompt}
conda activate renv

# install packages required
# pip install tqdm
# ...

conda deactivate
```

```{python}
# -*- coding: utf-8 -*-
"""
Created on Fri Jun  2 20:22:17 2023
Edited on Mon Dec 15 11:19:37 2025

@author: Wenhao
"""

import requests
import re
import os
import time
import pandas as pd
from tqdm import tqdm


os.chdir('//172.16.13.136/tphp/code/6_4drug')
df_cancer_drug = pd.read_excel('output/TPHP_cancer_drug_map_20251202.xlsx')

if not os.path.exists('tables_With/'): # output folder
    os.mkdir('tables_With')

for i, row in tqdm(df_cancer_drug.iterrows()):
    cancer = row['cancer']
    # drug = re.sub(r"[-(,\[\]\).']", " ", row["Name"])
    drug = row['Name'].replace('[', '(').replace(']', ')')
    # tb = f'https://clinicaltrials.gov/ct2/results/download_fields?down_count=10000&down_flds=all&down_fmt=csv&term={drug}&rslt=With&cond={cancer}&flds=a&flds=b&flds=e&flds=y'
    tb = f'https://clinicaltrials.gov/api/int/studies/download?format=csv&cond={cancer}&term={drug}&aggFilters=&sort=%40relevance&fields=NCT+Number%2CStudy+Title%2CStudy+URL%2CStudy+Status%2CConditions%2CInterventions%2CSponsor%2CCollaborators%2CStudy+Type%2CAcronym%2CBrief+Summary%2CStudy+Results%2CPrimary+Outcome+Measures%2CSecondary+Outcome+Measures%2COther+Outcome+Measures%2CSex%2CAge%2CPhases%2CEnrollment%2CFunder+Type%2CStudy+Design%2COther+IDs%2CStart+Date%2CPrimary+Completion+Date%2CCompletion+Date%2CFirst+Posted%2CResults+First+Posted%2CLast+Update+Posted%2CLocations%2CStudy+Documents'
    response = requests.get(tb)
    while response.status_code != 200:
        print(f'Error status code of {response.status_code}, sleep 5 sec...')
        time.sleep(5)
        response = requests.get(tb)
    print(response.status_code)
    content = response.content.decode(response.encoding).encode('utf-8')
    with open(f'tables_With/{cancer}_{drug}.csv', 'wb') as f:
        f.write(content)
```


## 1.3 combine tables
```{r}
tbls_cli <- list.files('tables_With/', full.names = T)

tbl_ls <- lapply(seq_along(tbls_cli), function(i){
  cat('Loading table ', i, ' of ', length(tbls_cli), '\n')
  fname <- tbls_cli[i]
  cancer <- fname %>% str_split('/') %>% unlist() %>% tail(1) %>%
    str_extract('.+?_') %>%
    str_replace('_', '') %>%
    str_replace('\\+', ' ')
  drug <- fname %>% str_split('/') %>% unlist() %>% tail(1) %>%
    str_extract('_.+?\\.csv') %>%
    str_replace('_', '') %>%
    str_replace('\\.csv', '')
  dftmp <- read.csv(fname, check.names = F)
  dftmp %<>% mutate(
    cancer = all_of(cancer),
    drug = all_of(drug)
  )
  return(dftmp)
})
df_cli <- plyr::ldply(tbl_ls)
df_cli %<>% select(`NCT Number`:drug)
df_cli %<>% filter(str_detect(`NCT Number`, '^NCT')) # remove unused rows

df_cli %>% rio::export('output/TPHP_dysregulate_clinical_trials_allResults_20251205.tsv')
df_cli %>% filter(`Study Results` == 'YES') %>% rio::export('output/TPHP_dysregulate_clinical_trials_WithResults_20251205.tsv')

df_cli %>% pull(`NCT Number`) %>% unique %>% length # 6868
df_cli %>% filter(`Study Results` == 'YES') %>% pull(`NCT Number`) %>% unique %>% length # 1703; 20230602version: 906

# nutrientional <- rio::import('//172.16.13.136/share/members/jiangwenhao/TPHP/20220908/clinical_trials_download/nutrientional_drugbank_results.csv') %>% pull(Name)
# df_cli %<>% filter(!(drug %in% nutrientional))
# df_cli %<>% filter(`Study Type` == 'Interventional')
# df_cli %<>% filter(Phases != 'Not Applicable')
# df_cli %>% rio::export('//172.16.13.136/share/members/jiangwenhao/TPHP/20220908/clinical_trials_download/TPHP_beast_endometrial_cancer_clinical_trials_hasResult_Interventional_phaseApplicable_without_nuterentional.tsv')
# 

df_drug_comb <- df_drug2 %>% rename(`Matched uniprot` = `Target uniprot`) %>% left_join(df_drug1)

df_cli_drug <- df_cli %>% left_join(df_drug_comb, c('drug' = 'Name'))
length(unique(df_cli_drug$`NCT Number`)) #6868; 20230602version: 3510
length(unique(df_cli_drug$`Matched uniprot`)) #290; 20230602version: 211
length(unique(df_cli_drug$drug)) #111; 20230602version: 70
df_cli_drug %>% rio::export('output/TPHP_DEP290_drug111_clinicalTrial6868_allResults_20251215.tsv')

df_cli_drug2 <- df_cli_drug %>% filter(!str_detect(drug, 'Insulin')) 
length(unique(df_cli_drug2$`NCT Number`)) #6086; 20230602version: 2089
length(unique(df_cli_drug2$`Matched uniprot`)) #284; 20230602version: 81
length(unique(df_cli_drug2$drug)) #108; 20230602version: 40


```

## 1.4 crawl clinical trials -- all results
### python crawler
```{python}
# # -*- coding: utf-8 -*-
# """
# Created on Fri Jun  2 20:22:17 2023
# 
# @author: Wenhao
# """
# 
# import requests
# import re
# import os
# import pandas as pd
# from tqdm import tqdm
# 
# 
# os.chdir('//172.16.13.136/share/members/jiangwenhao/TPHP/20220908/TPHP_carcinoma_upregulated_overlapped_drugbank')
# df_cancer_drug = pd.read_excel('//172.16.13.136/share/members/jiangwenhao/TPHP/20220908/TPHP_carcinoma_upregulated_overlapped_drugbank/TPHP_cancer_drug_map_20230602.xlsx')
# 
# if not os.path.exists('tables_all/'): # output folder
#     os.mkdir('tables_all')
# 
# for i, row in tqdm(df_cancer_drug.iterrows()):
#     cancer = row['cancer']
#     drug = row['Name']
#     tb = f'https://clinicaltrials.gov/ct2/results/download_fields?down_count=10000&down_flds=all&down_fmt=csv&term={drug}&rslt=all&cond={cancer}&flds=a&flds=b&flds=e&flds=y'
#     response = requests.get(tb)
#     content = response.content.decode(response.encoding).encode('utf-8')
#     with open(f'tables_all/{cancer}_{drug}.csv', 'wb') as f:
#         f.write(content)

```


## 1.5 combine tables -- all results
```{r}
# tbls_cli <- list.files('//172.16.13.136/share/members/jiangwenhao/TPHP/20220908/TPHP_carcinoma_upregulated_overlapped_drugbank/tables_all', full.names = T)
# 
# # 1-Azepan-1-Yl-2-Phenyl-2-(4-Thioxo-1,4-Dihydro-Pyrazolo[3,4-D]Pyrimidin-5-Yl)Ethanone Adduct
# # 语法错误，去掉
# tbls_cli <- tbls_cli[-1038]
# 
# 
# tbl_ls <- lapply(tbls_cli, function(fname){
#   cancer <- fname %>% str_split('/') %>% unlist() %>% tail(1) %>%
#     str_extract('.+?_') %>%
#     str_replace('_', '') %>%
#     str_replace('\\+', ' ')
#   drug <- fname %>% str_split('/') %>% unlist() %>% tail(1) %>%
#     str_extract('_.+?\\.csv') %>%
#     str_replace('_', '') %>%
#     str_replace('\\.csv', '')
#   dftmp <- read.csv(fname, check.names = F)
#   dftmp %<>% mutate(
#     cancer = all_of(cancer),
#     drug = all_of(drug)
#   )
#   return(dftmp)
# })
# df_cli <- plyr::ldply(tbl_ls)
# df_cli %<>% select(`NCT Number`:drug)
# df_cli %<>% filter(str_detect(`NCT Number`, '^NCT')) # remove unused rows
# 
# df_cli %>% rio::export('//172.16.13.136/share/members/jiangwenhao/TPHP/20220908/clinical_trials_download/TPHP_dysregulate_clinical_trials_allResults_20230602.tsv')
# 
# df_cli$`NCT Number` %>% unique %>% length # 3510
# 
# # nutrientional <- rio::import('//172.16.13.136/share/members/jiangwenhao/TPHP/20220908/clinical_trials_download/nutrientional_drugbank_results.csv') %>% pull(Name)
# # df_cli %<>% filter(!(drug %in% nutrientional))
# # df_cli %<>% filter(`Study Type` == 'Interventional')
# # df_cli %<>% filter(Phases != 'Not Applicable')
# # df_cli %>% rio::export('//172.16.13.136/share/members/jiangwenhao/TPHP/20220908/clinical_trials_download/TPHP_beast_endometrial_cancer_clinical_trials_hasResult_Interventional_phaseApplicable_without_nuterentional.tsv')
# # 
# 
# 
# df_drug_comb <- df_drug2 %>% rename(`Matched uniprot` = `Target uniprot`) %>% left_join(df_drug1)
# 
# df_cli_drug <- df_cli %>% left_join(df_drug_comb, c('drug' = 'Name'))
# length(unique(df_cli_drug$`NCT Number`)) # 3510
# length(unique(df_cli_drug$`Matched uniprot`)) # 211
# length(unique(df_cli_drug$drug)) # 70
# df_cli_drug %>% rio::export('//172.16.13.136/share/members/jiangwenhao/TPHP/20220908/clinical_trials_download/TPHP_DEP211_drug70_clinicalTrial3510_allResults_20230602.tsv')
# 

```

## 1.6 stat
```{r}
# df_cli_drug <- rio::import('//172.16.13.136/share/members/jiangwenhao/TPHP/20220908/clinical_trials_download/TPHP_DEP211_drug70_clinicalTrial3510_allResults_20230602.tsv')
df_cli_drug <- rio::import('output/TPHP_DEP290_drug111_clinicalTrial6868_allResults_20251215.tsv')

df_cli_drug %<>% mutate(`Last Update Posted` = as.character(`Last Update Posted`))
df_stat <- df_cli_drug %>%
  select(`NCT Number`, cancer, drug, Phases, contains('Date'), `Study Results`, `Target actions`, `Matched uniprot`, Approved:Withdrawn) %>%
  distinct()

dates <- df_stat %>% select(`Start Date`:`Last Update Posted`) %>% unlist() %>% unique() %>% sort()
# # change "December 6, 2019" to "2019-12-06"
# # This is not required now
# clinical_change_date <- function(x){
#   month2num <- sprintf("%02.f", 1:12)
#   names(month2num) <- month.name
#   x_split <- str_split(x, ' |, ')
#   sapply(x_split, function(xx) {
#     month <- month2num[xx[1]]
#     if(length(xx) == 2){
#       year <- xx[2]
#       ret <- str_c(year, month, sep = '-')
#     } else if(length(xx) == 3){
#       day <- sprintf("%02.f", as.numeric(xx[2]))
#       year <- xx[3]
#       ret <- str_c(year, month, day, sep = '-')
#     } else{
#       ret <- x
#     }
#     return(ret)
#   })
# }
# sum(dates != clinical_change_date(dates)) # 0
# df_stat %<>% mutate_at(which(colnames(.) == 'Start Date'):which(colnames(.) == 'Last Update Posted'), clinical_change_date)


df_stat$`Target actions` <- ifelse(df_stat$`Target actions` == '', 'Others', df_stat$`Target actions`)





# 以NCT number为中心汇总结果
df_stat_cli <- plyr::ddply(df_stat, '`NCT Number`', function(dfsub){
  cancers <- dfsub$cancer %>% unique %>% sort %>% str_c(collapse = '; ')
  drugs <- dfsub$drug %>% str_split('; ') %>% unlist %>% unique %>% sort %>% str_c(collapse = '; ')
  uniprots <- dfsub$`Matched uniprot` %>% str_split('; ') %>% unlist %>% unique %>% sort %>% str_c(collapse = '; ')
  actions <- dfsub$`Target actions` %>% str_split('; ') %>% unlist %>% unique %>% sort %>% str_c(collapse = '; ')
  status <- dfsub %>% select(Approved:Withdrawn) %>% apply(1, function(x) colnames(dfsub %>% select(Approved:Withdrawn))[x], simplify = F ) %>% unlist %>% unique %>% sort %>% str_c(collapse = '; ')
  df_ret <- dfsub %>% slice(1) %>% select(-(Approved:Withdrawn))
  df_ret$cancer <- cancers
  df_ret$drug <- drugs
  df_ret$`Matched uniprot` <- uniprots
  df_ret$`Target actions` <- actions
  df_ret$`Drug status` <- status
  return(df_ret)
})

# rio::export(df_stat_cli, 'output/TPHP_DEP290_drug111_clinicalTrial6868_allResults_stat_20251215.tsv')
df_stat_cli <- rio::import('output/TPHP_DEP290_drug111_clinicalTrial6868_allResults_stat_20251215.tsv')

df_drug_labeled <- df_cli_drug %>% distinct(`DrugBank ID`, drug, Description, Category, `Target actions`)
length(unique(df_drug_labeled$drug)) # 111
rio::export(df_drug_labeled, 'output/drugbank_results_all_20251215_111drugs.xlsx')

# df_drug %>% filter(`DrugBank ID` %in% c('DB00030', 'DB00005', 'DB06584', 'DB11731', 'DB12250', 'DB11840', 'DB05471', 'DB05101', 'DB06474', 'DB12845', 'DB05487', 'DB00089')) %>% rio::export('check1.tsv')

# After drugs munually classify：
df_drug_labeled_old <- readxl::read_excel('//172.16.13.136/share/members/jiangwenhao/TPHP/20220908/drug/drugbank_results_all_20230207_70drugs_tg(1).xlsx', sheet = 2)
df_drug_labeled <- rio::import('output/drugbank_results_all_20251215_111drugs.xlsx')


df_drug_labeled %>% count(Type3)

df_stat_cli1 <- df_stat_cli %>% inner_join(df_drug_labeled, by = c('drug' = 'Name'))
df_stat_cli1 %<>% filter(Phases != 'Not Applicable') # remove Not Applicable. Describes trials without FDA-defined phases, including trials of devices or behavioral interventions.
df_stat_cli1 %>% count(cancer)
```


### 1.6.2 circos

========================================
circlize version 0.4.15
CRAN page: https://cran.r-project.org/package=circlize
Github page: https://github.com/jokergoo/circlize
Documentation: https://jokergoo.github.io/circlize_book/book/

If you use it in published research, please cite:
Gu, Z. circlize implements and enhances circular visualization
  in R. Bioinformatics 2014.

This message can be suppressed by:
  suppressPackageStartupMessages(library(circlize))
========================================

```{r}
library(circlize)
df_stat_cli1 %>% count(Type3) # 5; drug types
df_stat_cli1 %>% count(`Last Update Posted`) # last date
df_stat_cli1[df_stat_cli1$Phases == 'Phase 1|Phase 2', 'Phases'] <- 'Phase 2'
df_stat_cli1[df_stat_cli1$Phases == 'Phase 2|Phase 3', 'Phases'] <- 'Phase 3'
df_stat_cli1[df_stat_cli1$Phases == '', 'Phases'] <- 'NA'
df_stat_cli1 %<>% filter(!str_detect(drug, 'Insulin')) # 去掉insulin相关
# df_stat_cli1 %>% rio::export('//172.16.13.136/share/members/jiangwenhao/TPHP/20220908/clinical_trials_download/TPHP_DEP81_drug40_clinicalTrial2089_allResults_excludingInsulin_stat_20230602.tsv')


# 2025.12.04 update
df_stat_cli1 <- rio::import('//172.16.13.136/share/members/jiangwenhao/TPHP/20220908/clinical_trials_download/TPHP_DEP81_drug40_clinicalTrial2089_allResults_excludingInsulin_stat_20230602.tsv')
prot_pattern <- str_c(sort(unique(df_dea1$protein[df_dea1$direction == 'Up'])), collapse = '|')
df_stat_cli1 %>% filter(!str_detect(`Matched uniprot`, prot_pattern)) %>%
  select(`NCT Number`, `Last Update Posted`, cancer, `Matched uniprot`, `Target actions`, `Drug status`, Phases, Type3) %>%
  rio::export('output/TPHP_DEP81_drug40_clinicalTrial2089_excluded_down_regulatioin.tsv')
df_stat_cli1 %<>% filter(str_detect(`Matched uniprot`, prot_pattern))
rio::export(df_stat_cli1, 'output/TPHP_DEP77_drug36_clinicalTrial2084_allResults_excludingInsulin_stat_20230602.tsv')
df_stat_cli1 %>%
  select(`NCT Number`, `Last Update Posted`, cancer, `Matched uniprot`, `Target actions`, `Drug status`, Phases, Type3) %>%
  rio::export('output/TPHP_DEP77_drug36_clinicaTrial2084_20230602_for_source_data.tsv')

df_plot <- df_stat_cli1 %>% select(`NCT Number`, `Last Update Posted`, cancer, `Matched uniprot`, `Target actions`, `Drug status`, Phases, Type3)



# 每圈是一个cancer type --- drug type 的label
cancer_types <- str_split(df_plot$cancer, '; ') %>% unlist %>% unique %>% sort
str_split(df_plot$cancer, '; ') %>% unlist %>% table %>% sort %>% rev
cancer_types_common <- c('Breast carcinoma', 'Ovarian carcinoma', 'Cervical carcinoma', 'Colon carcinoma', 'Hepatocellular carcinoma', 'Rectum carcinoma', 'Pancreas carcinoma', 'Endometrial carcinoma', 'Lung carcinoma')
df_plot$cancer <- str_replace_all(df_plot$cancer, setdiff(cancer_types, cancer_types_common) %>% str_c(collapse = '|'), 'Other cancers')
cancer_types <- str_split(df_plot$cancer, '; ') %>% unlist %>% unique %>% sort %>% .[. != 'Other cancers'] %>% c('Other cancers')

drug_types <- table(df_plot$Type3) %>% sort(decreasing = T) %>% names() %>% factor(., levels = ., ordered = T) # 按频次降序排列

# df_plot$cancer %<>% as.factor()
df_plot$`Last Update Posted` %<>% as.Date()
df_plot %<>% arrange(`Last Update Posted`, `NCT Number`)

# earliest <- as.numeric(min(df_plot$`Last Update Posted`))
# latest <- as.numeric(max(df_plot$`Last Update Posted`))
# dates <- earliest:latest

# circos
#prepare data
# df_color <- rio::import('Fig2_circle_scatter_col_xq_20200714.xlsx')
df_color <- data.frame(
  cancer_type = cancer_types,
  col_bg = c('#A079A588', '#E5C8B888', '#416F8888', '#AAAAB588', '#55BC6D88', '#D8894E88', '#ECC66D88', '#E5444488', '#9FB8DE88', '#8080BC88')
)

cd <- data.frame()
for(x1 in cancer_types){
  for(x2 in drug_types){
    cd %<>% rbind(data.frame(cancer_types = x1, drug_types = x2))
  }
}
cd$drug_types %<>% factor(., levels = levels(drug_types), ordered = T)
cd$cancer_types %<>% factor(., levels = unique(cd$cancer_types), ordered = T)
cd %<>% arrange(drug_types, cancer_types)

mat_candrug_NCT <- apply(df_plot, 1, function(x1){
  apply(cd, 1, function(x2){
    str_detect(x1['cancer'], x2['cancer_types']) & str_detect(x1['Type3'], x2['drug_types'])
  })
})
colnames(mat_candrug_NCT) <- df_plot$`NCT Number`
mat_candrug_NCT[mat_candrug_NCT] <- 1
mat_candrug_NCT[mat_candrug_NCT != 1] <- NA
df_candrug_NCT <- cbind(cd, mat_candrug_NCT)

# 删掉不包含的cancer-drug组合
row_remain <- rowSums(mat_candrug_NCT, na.rm = T) != 0
mat_candrug_NCT <- mat_candrug_NCT[row_remain, ]
df_candrug_NCT <- df_candrug_NCT[row_remain, ]

# 添加空白分隔行
last_row <- which(!duplicated(df_candrug_NCT$drug_types)) %>% setdiff(c(1))
for(i in rev(last_row)){ # 倒序插行
  df_candrug_NCT <- rbind(df_candrug_NCT[1:(i-1), ],
                          NA,
                          NA,
                          df_candrug_NCT[i:nrow(df_candrug_NCT), ])
}
rownames(df_candrug_NCT) <- NULL
mat_candrug_NCT <- df_candrug_NCT %>% select(-(1:2)) %>% as.matrix()

# df_color$pathway <- drug_types

# Start drawing (should be in console to be export)
circos.clear()
circos.par(cell.padding = c(0, 0, 0, 0), 
           start.degree = 90,
           gap.degree = 90,
           gap.after = 90,
           track.margin = c(0.001,0.001),
           track.height = 0.014)
circos.initialize(factors = factor(rep(1,length(df_plot$`NCT Number`))), 
                  x = 1:length(df_plot$`NCT Number`))
for(i in 1:nrow(df_candrug_NCT)){
  circos.trackPlotRegion(
    ylim = c(0, 2),
    track.index = i,
    bg.col = df_color$col_bg[df_color$cancer_type == df_candrug_NCT[i, 'cancer_types']], # 按照cancer types标背景颜色
    bg.border = NA, 
    panel.fun = function(x, y) {
      sector.index = get.cell.meta.data('sector.index')
      xlim = get.cell.meta.data('xlim')
      ylim = get.cell.meta.data('ylim')
    })
  
  y_vector <- mat_candrug_NCT[i, ]
  
  # y_up = y_vector
  # y_up[which(y_up < 0)] = NA
  # 
  # y_up_pos = y_up
  # y_up_pos[!is.na(y_up_pos)] = 1
  # 
  # y_down = y_vector
  # y_down[which(y_down > 0)] = NA  
  # 
  # y_down_pos = y_down
  # y_down_pos[!is.na(y_down_pos)] = 1
  # 
  # # Up_bg
  # circos.points(x = 1:length(gene5336),
  #               y = y_up_pos, 
  #               track.index = get.cell.meta.data("track.index"),
  #               col = "black",
  #               pch = 1,
  #               cex = abs(y_up)/5)
  # # Upregulation point
  # circos.points(track.index = get.cell.meta.data("track.index"),
  #               x = 1:length(gene5336),
  #               y = y_up_pos, 
  #               col = fig2_col$col_up[ceiling(i/10)],
  #               pch = 20,
  #               cex = abs(y_up)/5-0.02)
  # 
  # # Down_bg
  # circos.points(track.index = get.cell.meta.data("track.index"),
  #               x = 1:length(gene5336),
  #               y = y_down_pos, 
  #               col = "black",
  #               pch = 1,
  #               cex = abs(y_down)/5)
  # # Downregulation point
  # circos.points(track.index = get.cell.meta.data("track.index"),
  #               x = 1:length(gene5336),
  #               y = y_down_pos, 
  #               col = fig2_col$col_dw[ceiling(i/10)],
  #               pch = 20,
  #               cex = abs(y_down)/5-0.02)  
  
  # bg
  cex_bg <- sapply(df_plot$Phases, function(x) switch(x, 'NA' = 0.3, 'Early Phase 1' = 0.33, 'Phase 1' = 0.33, 'Phase 2' = 0.26, 'Phase 3' = 0.33, 'Phase 4' = 0.4)) * 3
  # circos.points(x = 1:length(df_plot$`NCT Number`),
  #               y = y_vector,
  #               track.index = get.cell.meta.data("track.index"),
  #               col = "black",
  #               pch = 1,
  #               cex = cex_bg)
  
  # point
  pch_point <- sapply(df_plot$Phases, function(x) switch(x, 'NA' = 3, 'Early Phase 1' = 20, 'Phase 1' = 20, 'Phase 2' = 17, 'Phase 3' = 18, 'Phase 4' = 8))
  col_point <- sapply(df_plot$Phases, function(x) switch(x, 'NA' = '#444444', 'Early Phase 1' = '#222222', 'Phase 1' = '#125789', 'Phase 2' = '#125789', 'Phase 3' = '#D22467', 'Phase 4' = '#D22467'))
  cex_point <- cex_bg - 0.02
  
  # names(col_point) <- names(pch_point) <- names(cex_point) <- names(y_vector)
  circos.points(track.index = get.cell.meta.data("track.index"),
                x = 1:length(df_plot$`NCT Number`),
                y = y_vector, 
                col = col_point,
                pch = pch_point,
                cex = cex_point)
}




# NCT labels
NCTs <- df_plot$`NCT Number`
names(NCTs) <- NCTs
NCTs_interested <- df_plot %>% filter(Phases == 'Phase 4') %>% pull(`NCT Number`) # Phase 4 trials
NCTs_mark <- rep(NA, length(NCTs))
names(NCTs_mark) <- NCTs
NCTs_mark[NCTs_interested] <- '-'
NCTs_mark_text <- NCTs_mark
# NCTs_mark_text[!is.na(NCTs_mark_text)] <- names(NCTs_mark_text[!is.na(NCTs_mark_text)])
NCTs_mark_text[!is.na(NCTs_mark_text)] <- df_stat_cli1 %>% filter(`NCT Number` %in% NCTs_interested) %>% arrange(`Last Update Posted`) %>% select(drug, cancer, `NCT Number`) %>% apply(1, function(x) str_glue("{x[1]}; {x[2]}; {x[3]}"))


#不需要加 “-”，直接修图拉直线
# circos.trackPlotRegion(factors = factor(rep(1,length(NCTs))),
#                        track.index = 1,
#                        panel.fun = function(x, y) {
#                          circos.text(1:length(NCTs),
#                                      CELL_META$ylim[1]+3,
#                                      NCTs_mark,
#                                      facing = "clockwise",
#                                      cex = 0.6,
#                                      niceFacing = TRUE,
#                                      adj = c(0, 0.5))
#                        },
#                        bg.border = NA)

circos.trackPlotRegion(factors = factor(rep(1,length(NCTs))),
                       track.index = 1,
                       panel.fun = function(x, y) {
                         circos.text(1:length(NCTs_mark),
                                     CELL_META$ylim[1]+7,
                                     NCTs_mark_text,
                                     facing = "clockwise",
                                     cex = 0.6,
                                     niceFacing = TRUE,
                                     adj = c(0, 0.5))
                       },
                       bg.border = NA)

# drug type labels
y_drug <- c(5, 17, 26, 34, 42, 49)  # seq(5,nrow(df_candrug_NCT),length(cancer_types))
for (j in 1:length(drug_types)) {
  circos.yaxis(side = "left",
               at = 1,
               track.index = y_drug[j],
               labels = drug_types[j],
               tick = F,
               sector.index = "1",
               labels.font = 0.8,
               col = "white")
}

#用颜色标注后不再需要文字labels
# # cancer type labels
# y_cancer <- 1:nrow(df_candrug_NCT)  # seq(5,nrow(df_candrug_NCT),length(cancer_types))
# for (j in 1:nrow(df_candrug_NCT)) {
#   circos.yaxis(side = "right",
#                at = 1,
#                track.index = y_cancer[j],
#                labels = df_candrug_NCT$cancer_types[j],
#                tick = F,
#                sector.index = "1",
#                labels.font = 0.8,
#                col = "white")
# }

# NCT date labels
x_date <- 1:ncol(mat_candrug_NCT)  # seq(5,nrow(df_candrug_NCT),length(cancer_types))
dates <- df_plot$`Last Update Posted`
years <- str_extract(dates, '^\\w+\\-') %>% str_replace('-', '')
min(years) # 2005
max(years) # 2023
# year_lbl_i <- which(!duplicated(years))
# years[year_lbl_i] <- 
years[duplicated(years)] <- NA

for (i in 1:ncol(mat_candrug_NCT)) {
  circos.axis(h = "bottom",
              sector.index = 1,
              track.index = nrow(mat_candrug_NCT),
              labels = as.character(years[i]),
              # tick = F,
              major.at = x_date[i],direction = "inside",
              labels.font = 0.8,labels.facing = "reverse.clockwise"
              #  col = "white"
  )
  
}


# add legend
lgd_links = ComplexHeatmap::Legend(
  title = 'Cancer name',
  at = df_color$cancer_type, # labels
  # col_fun = circlize::colorRamp2(seq_along(df_color$cancer_type), df_color$col_bg), # 连续型图例颜色
  legend_gp = grid::gpar(fill = df_color$col_bg), # 离散型图例颜色
  title_position = 'topleft'
)

ComplexHeatmap::draw(lgd_links, x = unit(1, "npc") - unit(2, "mm"), y = unit(4, "mm"), 
                     just = c("right", "bottom"))


# should save by manual
# 'TPHP_DEP_drug_clinicalTrial_20230626' # 9x9 size
'TPHP_DEP_drug_clinicalTrial_20251204' # 9x9 size

# circos.clear()




```

#### 1.7 discussion
```{r}
df_stat_cli1 %>% count(Phases)
# df_stat_cli1 %>% left_join(df_drug_labeled %>% select(Name, Type3) %>% distinct() %>% rename(drug = Name))
df_stat_cli1 %>% separate_rows(cancer, sep = '; ') %>% count(cancer, Type3) %>% arrange(desc(n))



X4 <- df_stat_cli1 %>% filter(Phases == 'Phase 4') %>% count(cancer, drug) %>% arrange(desc(n))
X3 <- df_stat_cli1 %>% filter(Phases == 'Phase 3') %>% count(cancer, drug) %>% arrange(desc(n))
X3 %>% filter(!(drug %in% X4$drug)) %>% arrange(desc(n))

df_drug_labeled %>% filter(Name == 'Amivantamab')


X0 <- df_stat_cli1 %>% filter(!(Phases %in% c('Phase 3', 'Phase 4'))) %>% count(cancer, drug) %>% arrange(desc(n))
X0 %>% filter(!(drug %in% c(X4$drug, X3$drug))) %>% arrange(desc(n))

# for manuscript
df_drug %>%
    separate_rows(`Target uniprot`, sep = ';') %>%
    filter(`Target uniprot` != '') %>% pull(`DrugBank ID`) %>% unique() %>% length() # 8179
df_drug %>%
    separate_rows(`Target uniprot`, sep = ';') %>%
    filter(`Target uniprot` != '') %>% pull(`Target uniprot`) %>% unique() %>% length() # 6653
length(unique(df_drug2$`DrugBank ID`)) # 528
length(unique(df_drug2$`Target uniprot`)) # 573

length(intersect(df_dea$protein, df_drug2$`Target uniprot`)) # 301
length(intersect(df_drug2$`Target uniprot`, df_dea1$protein)) # 260


# length(unique(df_stat_cli1$drug)) # 40
# length(unique(df_stat_cli1$`NCT Number`)) # 2089
# length(unique(df_plot$`NCT Number`)) # 2089
# length(unique(unlist(str_split(df_plot$`Matched uniprot`, '; ')))) # 81
# length(unique(unlist(str_split(df_plot$cancer, '; ')))) # 10


length(unique(unlist(str_split(df_stat_cli1$`Matched uniprot`, '; ')))) # 77
length(unique(df_stat_cli1$drug)) # 36
length(unique(df_stat_cli1$`NCT Number`)) # 2084

# We downloaded all DrugBank results (8,179 drugs targeting 6,653 proteins; 2023-02-07) and then restricted the dataset to biotech drugs (528 drugs targeting 573 proteins). Of these targets, 301 were quantified in our paired tumor dysregulation analysis, and 260 were identified as DEPs. And 40 of these drugs with DEP-corresponding cancers matched to clinicalTrial database, corresponding to 81 DEPs in 10 cancer types. 

# We downloaded all DrugBank results (8179 drugs targeting 6653 proteins; 2023-02-07) and then focused on biotech drugs (528 drugs targeting 573 proteins). Among these target proteins, 301 were quantified in our paired tumor dysregulation analysis, and 260 were identified as DEPs. After matching the drugs to ClinicalTrials.gov database by cancer type, 40 drugs had clinical trial records in the corresponding cancers, involving 81 DEPs across 10 cancer types.

```



# ===================================
# 1. drugbank, clinical trials
```{r}
tbls <- read_excel_allsheets('//172.16.13.136/share/members/jiangwenhao/TPHP/20220908/TPHP_carcinoma_upregulated_overlapped_drugbank/TPHP_carcinoma_upregulated_overlapped_target_drugbank_20220926.xlsx')
db_ids <- lapply(tbls, function(df){
  if('drug' %in% colnames(df))
    return(unique(df$drug))
  if('DrugBank ID' %in% colnames(df))
    return(unique(df$`DrugBank ID`))
})

db_id_unique <- unique(unlist(db_ids))
df_cli_ls <- list()
for (i in 1:length(db_id_unique)) {
  print(str_glue('Search drugid {i}/{length(db_id_unique)}'))
  drug_id <- db_id_unique[i]
  df_cli_ls[[i]] <- clinical_trial_of_drug(drug_id)
}

names(df_cli_ls) <- db_id_unique
for(i in seq_along(df_cli_ls)){
  if(nrow(df_cli_ls[[i]]) == 0){
    df_cli_ls[[i]] <- data.frame(
      `DrugBank ID` = NA,
      PAHSE = NA,
      STATUS = NA,
      PURPOSE = NA,
      CONDITIONS = NA,
      CONDITIONS_LINK = NA,
      COUNT = NA,
      COUNT_LINK = NA,
      check.names = F
    )
  }
}
df_cli_all <- plyr::ldply(df_cli_ls, .id = 'DrugBankID')
rio::export(df_cli_all, '//172.16.13.136/share/members/jiangwenhao/TPHP/20220908/TPHP_carcinoma_upregulated_overlapped_drugbank/TPHP_carcinoma_upregulated_proteins_drugbank_clinical_trials_3754.xlsx')



df_cli_c <- df_cli_all %>%
  filter(str_detect(str_to_lower(CONDITIONS), 'leukemia|cancer|tumor|[a-z]+oma[^a-z]'))
df_cli_c$CONDITIONS %<>% str_replace_all('&#39;s', "'s")

df_cli_c <- rio::import('//172.16.13.136/share/members/jiangwenhao/TPHP/20220908/TPHP_carcinoma_upregulated_overlapped_drugbank/TPHP_carcinoma_upregulated_proteins_drugbank_clinical_trials_3754_cancertumoromaleukemia.xlsx')

indice <- str_which(df_cli_c$CONDITIONS, '<a href.+?>.+?;.+?</a>')
for(i in indice){
  s <- df_cli_c$CONDITIONS[i]
  s_replaced_vec <- str_extract_all(s, '<a href.+?>.+?;.+?</a>')[[1]]
  for(s_from in s_replaced_vec){
    s_to <- str_match_all(s_from, '>(.+?)</a>')[[1]][1, 2]
    s_to %<>% gsub(pattern = ';', replacement = '；', ., fixed = T)
    s %<>% gsub(pattern = s_from, replacement = s_to, ., fixed = T)
  }
  
  df_cli_c[i, 'CONDITIONS'] <- s
}

rio::export(df_cli_c, '//172.16.13.136/share/members/jiangwenhao/TPHP/20220908/TPHP_carcinoma_upregulated_overlapped_drugbank/TPHP_carcinoma_upregulated_proteins_drugbank_clinical_trials_3754_cancertumoromaleukemia.xlsx')


```



## conditions
```{r}
df_cli_c <- rio::import('//172.16.13.136/share/members/jiangwenhao/TPHP/20220908/TPHP_carcinoma_upregulated_overlapped_drugbank/TPHP_carcinoma_upregulated_proteins_drugbank_clinical_trials_3754_cancertumoromaleukemia.xlsx')

conditions <- unique(df_cli_c$CONDITIONS) %>%
  # str_split(';') %>% unlist() %>% unique() %>%
  sort()
data.frame(condition = conditions) %>% rio::export('conditions.xlsx')



df_cli_ct <- df_cli_c %>%
  filter(str_detect(str_to_lower(CONDITIONS), 'cancer|tumor'))


df_cli_ct1 <- df_cli_ct %>%
  filter(str_detect(CONDITIONS, '[Bb]reast|[Ee]ndometrial')) %>%
  filter(PURPOSE == 'Treatment') %>%
  arrange(DrugBankID, PAHSE, STATUS)
df_cli_ct1 %>%
  rio::export('//172.16.13.136/share/members/jiangwenhao/TPHP/20220908/TPHP_carcinoma_upregulated_overlapped_drugbank/TPHP_carcinoma_upregulated_proteins_drugbank_clinical_trials_3754_cancertumor_breastEndometrial.xlsx')


dbid_selected <- sapply(unique(df_cli_ct1$DrugBankID), function(x){
  dfsub <- df_cli_ct1 %>% filter(DrugBankID == all_of(x))
  if(any(str_detect(dfsub$CONDITIONS, '[Bb]reast')) & any(str_detect(dfsub$CONDITIONS, '[Ee]ndometrial'))){
    ;
  } else{
    x <- NA
  }
  return(x)
}) %>% .[!is.na(.)]

df_cli_ct2 <- df_cli_ct1 %>% filter(DrugBankID %in% dbid_selected)
df_cli_ct2 %>%
  rio::export('//172.16.13.136/share/members/jiangwenhao/TPHP/20220908/TPHP_carcinoma_upregulated_overlapped_drugbank/TPHP_carcinoma_upregulated_proteins_drugbank_clinical_trials_3754_cancertumor_breastEndometrial_both_20221011.xlsx')


df_tbls <- plyr::ldply(tbls, function(df){
  if('drug' %in% colnames(df))
    return(df %>% rename(`DrugBank ID` = drug))
  if('DrugBank ID' %in% colnames(df))
    return(df)
}, .id = 'drug_stage')

df_tbls1 <- df_tbls %>%
  select(`DrugBank ID`, drug_stage, Name, drug_stage, organ, UniprotID, Description, `Target name`, `Target actions`, `Target uniprot`) %>%
  distinct() %>%
  filter(drug_stage %in% c('approved', 'illicit', 'investigatioinal', 'experimental', 'chemotherapy')) %>%
  filter(`DrugBank ID` %in% dbid_selected) %>%
  rename(DrugBankID = `DrugBank ID`) %>%
  group_by(DrugBankID) %>%
  summarise(
    Name = str_c(sort(unique(Name)), collapse = ';'),
    drug_status = str_c(sort(unique(drug_stage)), collapse = ';'),
    organ = str_c(sort(unique(organ)), collapse = ';'),
    UniprotID = str_c(sort(unique(UniprotID)), collapse = ';'),
    Description = str_c(sort(unique(Description)), collapse = ';'),
    `Target name` = str_c(sort(unique(`Target name`)), collapse = ';'),
    `Target actions` = str_c(sort(unique(`Target actions`)), collapse = ';'),
    `Target uniprot` = str_c(sort(unique(`Target uniprot`)), collapse = ';')
  )

df_tbls1 %<>% filter(str_detect(organ, 'mammary'), str_detect(organ, 'uterus')) # mammary uterus overlapped


df_cli_ct3 <- inner_join(df_cli_ct2, df_tbls1, by = 'DrugBankID')
df_cli_ct3 <- df_cli_ct3[, -which(colnames(df_cli_ct3) == 'DrugBank ID')]

# df_cli_ct3 %>%
#   rio::export('//172.16.13.136/share/members/jiangwenhao/TPHP/20220908/TPHP_carcinoma_upregulated_overlapped_drugbank/TPHP_carcinoma_upregulated_proteins_drugbank_clinical_trials_3754_cancertumor_breastEndometrial_both_drugstatus_target.xlsx')
# 

df_cli_ct3 %<>% filter(str_detect(organ, 'mammary'), str_detect(organ, 'uterus')) # select TPHP mammary gland and uterus overlapped

df_cli_ct3 %>%
  rio::export('//172.16.13.136/share/members/jiangwenhao/TPHP/20220908/TPHP_carcinoma_upregulated_overlapped_drugbank/TPHP_carcinoma_upregulated_proteins_drugbank_clinical_trials_3754_cancertumor_breastEndometrial_only_both_drugstatus_target_20221011.xlsx')





# drugs targeted only one organ
dbid_selected <- sapply(unique(df_cli_ct1$DrugBankID), function(x){
  dfsub <- df_cli_ct1 %>% filter(DrugBankID == all_of(x))
  if(xor(any(str_detect(dfsub$CONDITIONS, '[Bb]reast')), any(str_detect(dfsub$CONDITIONS, '[Ee]ndometrial')))){ # exclusive OR
    ;
  } else{
    x <- NA
  }
  return(x)
}) %>% .[!is.na(.)]

df_cli_ct2 <- df_cli_ct1 %>% filter(DrugBankID %in% dbid_selected)
# df_cli_ct2 %>%
#   rio::export('//172.16.13.136/share/members/jiangwenhao/TPHP/20220908/TPHP_carcinoma_upregulated_overlapped_drugbank/TPHP_carcinoma_upregulated_proteins_drugbank_clinical_trials_3754_cancertumor_breastEndometrial_both_20221011.xlsx')


df_tbls <- plyr::ldply(tbls, function(df){
  if('drug' %in% colnames(df))
    return(df %>% rename(`DrugBank ID` = drug))
  if('DrugBank ID' %in% colnames(df))
    return(df)
}, .id = 'drug_stage')

df_tbls1 <- df_tbls %>%
  select(`DrugBank ID`, drug_stage, Name, drug_stage, organ, UniprotID, Description, `Target name`, `Target actions`, `Target uniprot`) %>%
  distinct() %>%
  filter(drug_stage %in% c('approved', 'illicit', 'investigatioinal', 'experimental', 'chemotherapy')) %>%
  filter(`DrugBank ID` %in% dbid_selected) %>%
  rename(DrugBankID = `DrugBank ID`) %>%
  group_by(DrugBankID) %>%
  summarise(
    Name = str_c(sort(unique(Name)), collapse = ';'),
    drug_status = str_c(sort(unique(drug_stage)), collapse = ';'),
    organ = str_c(sort(unique(organ)), collapse = ';'),
    UniprotID = str_c(sort(unique(UniprotID)), collapse = ';'),
    Description = str_c(sort(unique(Description)), collapse = ';'),
    `Target name` = str_c(sort(unique(`Target name`)), collapse = ';'),
    `Target actions` = str_c(sort(unique(`Target actions`)), collapse = ';'),
    `Target uniprot` = str_c(sort(unique(`Target uniprot`)), collapse = ';')
  )

df_tbls1 %<>% filter(str_detect(organ, 'mammary'), str_detect(organ, 'uterus')) # mammary uterus overlapped


df_cli_ct3 <- inner_join(df_cli_ct2, df_tbls1, by = 'DrugBankID')
df_cli_ct3 <- df_cli_ct3[, -which(colnames(df_cli_ct3) == 'DrugBank ID')]

# df_cli_ct3 %>%
#   rio::export('//172.16.13.136/share/members/jiangwenhao/TPHP/20220908/TPHP_carcinoma_upregulated_overlapped_drugbank/TPHP_carcinoma_upregulated_proteins_drugbank_clinical_trials_3754_cancertumor_breastEndometrial_both_drugstatus_target.xlsx')
# 

df_cli_ct3 %<>% filter(str_detect(organ, 'mammary'), str_detect(organ, 'uterus')) # select TPHP mammary gland and uterus overlapped

df_cli_ct3 %>%
  rio::export('//172.16.13.136/share/members/jiangwenhao/TPHP/20220908/TPHP_carcinoma_upregulated_overlapped_drugbank/TPHP_carcinoma_upregulated_proteins_drugbank_clinical_trials_3754_cancertumor_breastEndometrial_only_one_drugstatus_target_20221011.xlsx')


```


## breast cancer and endometrial cancer
```{r}
df_cli_2 <- rio::import('//172.16.13.136/share/members/jiangwenhao/TPHP/20220908/TPHP_carcinoma_upregulated_overlapped_drugbank/TPHP_carcinoma_upregulated_proteins_drugbank_clinical_trials_3754_cancertumor_breastEndometrial_only_both_drugstatus_target_20221011.xlsx')
df_cli_1 <- rio::import('//172.16.13.136/share/members/jiangwenhao/TPHP/20220908/TPHP_carcinoma_upregulated_overlapped_drugbank/TPHP_carcinoma_upregulated_proteins_drugbank_clinical_trials_3754_cancertumor_breastEndometrial_only_one_drugstatus_target_20221011.xlsx')



```


### crawl clinical trials
```{python}
import requests
import re
import os
import pandas as pd
from tqdm import tqdm


# main
df_cli_1 = pd.read_excel('//172.16.13.136/share/members/jiangwenhao/TPHP/20220908/TPHP_carcinoma_upregulated_overlapped_drugbank/TPHP_carcinoma_upregulated_proteins_drugbank_clinical_trials_3754_cancertumor_breastEndometrial_only_one_drugstatus_target_20221011.xlsx')
df_cli_2 = pd.read_excel('//172.16.13.136/share/members/jiangwenhao/TPHP/20220908/TPHP_carcinoma_upregulated_overlapped_drugbank/TPHP_carcinoma_upregulated_proteins_drugbank_clinical_trials_3754_cancertumor_breastEndometrial_only_both_drugstatus_target_20221011.xlsx')


drugs = sorted(list(set(df_cli_1['Name']) | set(df_cli_2['Name'])))
cancers = ['Endometrial+Cancer', 'Breast+Cancer']
if not os.path.exists('//172.16.13.136/share/members/jiangwenhao/TPHP/20220908/clinical_trials_download/tables/'): # output folder
    os.mkdir('//172.16.13.136/share/members/jiangwenhao/TPHP/20220908/clinical_trials_download/tables')
for cancer in cancers:
    for drug in tqdm(drugs):
        # url = f'https://clinicaltrials.gov/ct2/results?type=&rslt=With&cond={cancer}&term={drug}'
        tb = f'https://clinicaltrials.gov/ct2/results/download_fields?down_count=10000&down_flds=all&down_fmt=csv&term={drug}&rslt=With&cond={cancer}&flds=a&flds=b&flds=e&flds=y'
        response = requests.get(tb)
        content = response.content.decode(response.encoding).encode('utf-8')
        with open(f'//172.16.13.136/share/members/jiangwenhao/TPHP/20220908/clinical_trials_download/tables/{cancer}_{drug}.csv', 'wb') as f:
            f.write(content)


```



### combine tables
```{r}
tbls_cli <- list.files('//172.16.13.136/share/members/jiangwenhao/TPHP/20220908/clinical_trials_download/tables', full.names = T)
tbl_ls <- lapply(tbls_cli, function(fname){
  cancer <- fname %>% str_split('/') %>% unlist() %>% tail(1) %>%
    str_extract('.+?_') %>%
    str_replace('_', '') %>%
    str_replace('\\+', ' ')
  drug <- fname %>% str_split('/') %>% unlist() %>% tail(1) %>%
    str_extract('_.+?\\.csv') %>%
    str_replace('_', '') %>%
    str_replace('\\.csv', '')
  dftmp <- read.csv(fname, check.names = F)
  dftmp %<>% mutate(
    cancer = all_of(cancer),
    drug = all_of(drug)
  )
  return(dftmp)
})
df_cli <- plyr::ldply(tbl_ls)
df_cli %<>% select(`NCT Number`:drug)
df_cli %<>% filter(str_detect(`NCT Number`, '^NCT')) # remove unused rows
# df_cli %>% rio::export('//172.16.13.136/share/members/jiangwenhao/TPHP/20220908/clinical_trials_download/TPHP_beast_endometrial_cancer_clinical_trials_hasResult.csv')
df_cli %>% rio::export('//172.16.13.136/share/members/jiangwenhao/TPHP/20220908/clinical_trials_download/TPHP_beast_endometrial_cancer_clinical_trials_hasResult.tsv')
df_cli <- rio::import('//172.16.13.136/share/members/jiangwenhao/TPHP/20220908/clinical_trials_download/TPHP_beast_endometrial_cancer_clinical_trials_hasResult.tsv')

nutrientional <- rio::import('//172.16.13.136/share/members/jiangwenhao/TPHP/20220908/clinical_trials_download/nutrientional_drugbank_results.csv') %>% pull(Name)
df_cli %<>% filter(!(drug %in% nutrientional))
df_cli %<>% filter(`Study Type` == 'Interventional')
df_cli %<>% filter(Phases != 'Not Applicable')
df_cli %>% rio::export('//172.16.13.136/share/members/jiangwenhao/TPHP/20220908/clinical_trials_download/TPHP_beast_endometrial_cancer_clinical_trials_hasResult_Interventional_phaseApplicable_without_nuterentional.tsv')


```



### stat: protein, drug, phase, cancer
```{r}
df_cli_simple <- df_cli %>% select(drug, cancer, Conditions, Phases) %>% distinct()
df_drug_prot <- rbind(df_cli_1, df_cli_2) %>%
  select(Name, organ, UniprotID) %>%
  distinct()

df_cli_simple %<>% inner_join(df_drug_prot, by = c('drug' = 'Name'))
df_cli_simple %>% rio::export('//172.16.13.136/share/members/jiangwenhao/TPHP/20220908/clinical_trials_download/TPHP_beast_endometrial_cancer_simple.tsv')

df_cli_simple1 <- df_cli_simple %>% select(UniprotID, drug, Phases, cancer) %>%
  separate_rows(UniprotID) %>%
  distinct() %>%
  arrange(UniprotID, drug, Phases, cancer)
df_cli_simple1 %>% rio::export('//172.16.13.136/share/members/jiangwenhao/TPHP/20220908/clinical_trials_download/TPHP_beast_endometrial_cancer_clinical_trials_355proteins_119drugs.xlsx')

```

### search drug
```{r}
# df_cli_simple1 <- rio::import('//172.16.13.136/share/members/jiangwenhao/TPHP/20220908/clinical_trials_download/TPHP_beast_endometrial_cancer_clinical_trials_355proteins_119drugs.xlsx')
# drugs <- unique(df_cli_simple1$drug)
# df_drug_ls <- list()
# for (i in 1:length(drugs)) {
#   print(str_glue('Search drug {drugs[i]}    {i}/{length(drugs)}'))
#   drug_name <- drugs[i]
#   df_drug_ls[[i]] <- drugbank_search_name(drug_name)
# }
# df_drug <- plyr::ldply(df_drug_ls)
# df_drug %>% rio::export('//172.16.13.136/share/members/jiangwenhao/TPHP/20220908/clinical_trials_download/drugs119_information.xlsx')


# filter manually
df_cli1 <- rio::import('//172.16.13.136/share/members/jiangwenhao/TPHP/20220908/clinical_trials_download/TPHP_beast_endometrial_cancer_clinical_trials_hasResult_Interventional_phaseApplicable_without_nuterentional.tsv')
df_drug <- rio::import('//172.16.13.136/share/members/jiangwenhao/TPHP/20220908/clinical_trials_download/drugs119_information.xlsx')

drug_used <- df_drug %>% filter(Removed == 'F') %>% pull(GenericName) %>% unique()
df_cli2 <- df_cli1 %>% filter(drug %in% drug_used)
df_cli2 %>% rio::export('//172.16.13.136/share/members/jiangwenhao/TPHP/20220908/clinical_trials_download/TPHP_beast_endometrial_cancer_clinical_trials_hasResult_Interventional_phaseApplicable_without_nuterentional_antitumorDrug.tsv')



df_cli2$`NCT Number` %>% unique() %>% length() # 748
df_cli2 %>% select(-drug, -cancer) %>% distinct() %>% nrow() # 748
# NCT_tmps <- df_cli2 %>% select(-drug, -cancer) %>% distinct() %>% count(`NCT Number`) %>% filter(n > 1) %>% pull(`NCT Number`)
# df_cli2 %>% filter(`NCT Number` %in% NCT_tmps) %>% arrange(`NCT Number`) %>% View()
# NCT_multidrug <- df_cli2 %>% count(`NCT Number`) %>% filter(n > 1) %>% pull(`NCT Number`)
# df_cli2 %>% filter(`NCT Number` %in% NCT_multidrug) %>% arrange(`NCT Number`)

colnames(df_cli2)[1] <- 'NCTNumber'
df_cli3 <- plyr::ddply(df_cli2, .variables = 'NCTNumber', function(dfsub){
  dfret <- dfsub %>% select(-drug, -cancer) %>% distinct()
  dfret$cancer <- dfsub$cancer %>% unique() %>% sort() %>% str_c(collapse = ';')
  dfret$drug <- dfsub$drug %>% unique() %>% sort() %>% str_c(collapse = ';')
  return(dfret)
})
colnames(df_cli3)[1] <- colnames(df_cli2)[1] <- 'NCT Number'
df_cli3 %>% rio::export('//172.16.13.136/share/members/jiangwenhao/TPHP/20220908/clinical_trials_download/TPHP_beast_endometrial_cancer_clinical_trials_hasResult_Interventional_phaseApplicable_without_nuterentional_antitumorDrug_748trials.tsv')
df_cli3$`NCT Number` %>% unique() %>% length() # 748

df_cli3_simple <- df_cli3 %>%
  select(`NCT Number`, Title, Status, Conditions, Phases, cancer, drug) %>%
  arrange(Phases, cancer)

df_cli3_simple %>%
  rio::export('//172.16.13.136/share/members/jiangwenhao/TPHP/20220908/clinical_trials_download/748trials_simple.xlsx')


df_cli3_simple %>% count(cancer, Phases)
df_cli3_simple %>% filter(!str_detect(df_cli3_simple$drug, ';')) %>%
  count(drug, Phases)

```


### drugs分类 (根据targets分类)
```{r}
df_cli_simple1 <- rio::import('//172.16.13.136/share/members/jiangwenhao/TPHP/20220908/clinical_trials_download/TPHP_beast_endometrial_cancer_clinical_trials_355proteins_119drugs.xlsx')

df_cli3 <- rio::import('//172.16.13.136/share/members/jiangwenhao/TPHP/20220908/clinical_trials_download/TPHP_beast_endometrial_cancer_clinical_trials_hasResult_Interventional_phaseApplicable_without_nuterentional_antitumorDrug_748trials.tsv')


df_drug_prot <- df_cli_simple1 %>% distinct(UniprotID, drug) # UniprotID, drug

# immunotherapy targets
df_immu <- readxl::read_excel('//172.16.13.136/share/members/jiangwenhao/TPHP/20220908/clinical_trials_download/drugbank_result_immune_therapy_combined_edited2_reviewed_v2_sheet4_20221026.xlsx') %>% select(DRUG, TARGET)
targets_immu <- unique(df_immu$TARGET) %>% .[!is.na(.)]


# rtk targets
df_rtk <- readxl::read_excel('//172.16.13.136/share/members/jiangwenhao/TPHP/20220908/clinical_trials_download/RTKs.xlsx')
uniprots_rtk <- unique(df_rtk$From) %>% .[!is.na(.)]


# 癌症激素治療
df_horm1 <- readxl::read_excel('//172.16.13.136/share/members/jiangwenhao/TPHP/20220908/clinical_trials_download/drugbank_results_hormoral.xlsx', sheet = 'drugbank_results_hormoral')
df_horm2 <- readxl::read_excel('//172.16.13.136/share/members/jiangwenhao/TPHP/20220908/clinical_trials_download/drugbank_results_hormoral.xlsx', sheet = 'Catogarized targets')
uniprots_horm1 <- unique(df_horm1$`Target uniprot`) %>%
  .[!is.na(.)] %>%
  str_split(';') %>%
  unlist() %>%
  unique()
targets_horm2 <- unique(df_horm2$Target) %>% .[!is.na(.)]




targets <- union(targets_immu, targets_horm2)
# uniprots <- c()
# for(i in seq_along(targets)){
#   print(str_glue('Search target {i}/{length(targets)}'))
#   # uniprots[i] <- drugbank_search_protname(targets[i])
#   uniprots[i] <- protname_to_uniprotid(targets[i])
# }
# df_target_uniprot <- data.frame(target = targets, uniprotid = uniprots)
# df_target_uniprot %>% rio::export('//172.16.13.136/share/members/jiangwenhao/TPHP/20220908/clinical_trials_download/243immn_horm_targets_uniprot.xlsx')
df_target_uniprot <- rio::import('//172.16.13.136/share/members/jiangwenhao/TPHP/20220908/clinical_trials_download/243immn_horm_targets_uniprot_edited.xlsx')
df_target_uniprot_sep <- df_target_uniprot %>% separate_rows('uniprotid', sep = ';') %>%
  filter(uniprotid != '-')
ht_target_uniprot <- df_target_uniprot_sep$uniprotid
names(ht_target_uniprot) <- df_target_uniprot_sep$target


df1 <- data.frame(target = targets_immu, uniprotid = ht_target_uniprot[targets_immu] %>% unname(), immuno = T, rtk = NA, hormoral = NA)
df2 <- data.frame(target = NA, uniprotid = uniprots_rtk, immuno = NA, rtk = T, hormoral = NA)
df3 <- data.frame(target = NA, uniprotid = uniprots_horm1, immuno = NA, rtk = NA, hormoral = T)
df4 <- data.frame(target = targets_horm2, uniprotid = ht_target_uniprot[targets_horm2] %>% unname(), immuno = NA, rtk = NA, hormoral = T)

# Reduce(function(x, y){
#   full_join(x, y, by = 'uniprotid')
# }, list(df1, df2, df3, df4))
Reduce(intersect, list(df1$uniprotid, df2$uniprotid, df3$uniprotid, df4$uniprotid))
df <- rbind(df1, df2, df3, df4)
df[!is.na(df$immuno), 'immuno'] <- 'immuno'
df[!is.na(df$rtk), 'rtk'] <- 'rtk'
df[!is.na(df$hormoral), 'hormoral'] <- 'hormoral'
length(unique(df$uniprotid)) # 3776
df$uniprotid %<>% str_trim()
length(unique(df$uniprotid)) # 3109

df_target <- df %>% drop_na(uniprotid) %>%
  filter(uniprotid != '-') %>%
  filter(uniprotid != '') %>%
  pivot_longer(cols = c('immuno', 'rtk', 'hormoral')) %>%
  drop_na(value) %>%
  select(-name) %>%
  distinct() %>%
  pivot_wider(id_cols = c('target', 'uniprotid'), names_from = value)

df_target %<>%
  mutate(
    immuno = ifelse(!is.na(immuno), T, F),
    rtk = ifelse(!is.na(rtk), T, F),
    hormoral = ifelse(!is.na(hormoral), T, F)
  )
df_target %>% rio::export('//172.16.13.136/share/members/jiangwenhao/TPHP/20220908/clinical_trials_download/drug_target_classification.xlsx')

length(unique(df_target$uniprotid)) # 3107



uniprotid_dup <- df_target$uniprotid %>% .[duplicated(.)]
df_target_dup <- df_target %>%
  filter(uniprotid %in% uniprotid_dup) %>%
  arrange(uniprotid, target)
rio::export(df_target_dup, '//172.16.13.136/share/members/jiangwenhao/TPHP/20220908/clinical_trials_download/drug_target_classification_dup444.xlsx')

df_target_dup %>%
  count(uniprotid) %>%
  filter(n > 2)

df_target_dup_combine <- df_target_dup %>%
  plyr::ddply('uniprotid', function(dfsub){
    dfret <- dfsub[1, ]
    dfret$target <- dfsub$target %>% unique() %>% .[!is.na(.)] %>% str_c(collapse = ';')
    dfret$immuno <- any(dfsub$immuno)
    dfret$rtk <- any(dfsub$rtk)
    dfret$hormoral <- any(dfsub$hormoral)
    return(dfret)
  })

df_target_rlt <- df_target %>%
  filter(!(uniprotid %in% uniprotid_dup)) %>%
  rbind(df_target_dup_combine)
  arrange(uniprotid)
rio::export(df_target_rlt, '//172.16.13.136/share/members/jiangwenhao/TPHP/20220908/clinical_trials_download/drug_target_classification_3107uniqueUniprotID.xlsx')



df_cli_simple1_ <- df_cli_simple1 %>%
  left_join(df_target_rlt, by = c('UniprotID' = 'uniprotid'))

rio::export(df_cli_simple1_, '//172.16.13.136/share/members/jiangwenhao/TPHP/20220908/clinical_trials_download/clinical_trials_target_classification.xlsx')

# df_cli3 %>% left_join(df_cli_simple1_, by = 'drug')
df_cli_simple1_$drug %>% setdiff(unlist(str_split(df_cli3$drug, ';')))
df_cli_simple1_$drug %>% setdiff(unlist(str_split(df_cli3$drug, ';')), .) # character(0)

df_target_uniprot_drug <- df_target_rlt %>%
  left_join(df_cli_simple1, by = c('uniprotid' = 'UniprotID')) %>%
  select(-Phases, -cancer) %>%
  distinct()
rio::export(df_target_uniprot_drug, '//172.16.13.136/share/members/jiangwenhao/TPHP/20220908/clinical_trials_download/target_uniprotid_drug.xlsx')

drug_types_of_trial <- lapply(df_cli3$drug, function(drugs){
  drugs %<>% str_split(';') %>% unlist()
  ret <- df_target_uniprot_drug %>%
    filter(drug %in% drugs) %>%
    summarise_at(vars(immuno, rtk, hormoral), sum) %>%
    mutate_all(function(x) ifelse(x > 0, T, F)) %>%
    mutate(others = ifelse(length(setdiff(drugs, df_target_uniprot_drug$drug)) > 0, T, F))
  return(ret)
})
df_drug_label <- Reduce(rbind, drug_types_of_trial) %>%
  add_column(drug = df_cli3$drug, .before = 1)
df_drug_label %>% select(-drug) %>%
  cbind(df_cli3, .)

df_cli <- df_drug_label %>% select(-drug) %>%
    cbind(df_cli3, .)

df_cli %>% count(immuno)
df_cli %>% count(rtk)
df_cli %>% count(hormoral)

df_drug_types <- df_cli %>% select(`NCT Number`, drug:others) %>%
  mutate(
    immuno = ifelse(immuno, 'immuno', NA),
    rtk = ifelse(rtk, 'rtk', NA),
    hormoral = ifelse(hormoral, 'hormoral', NA),
    others = ifelse(others, 'others', NA)
  ) %>%
  pivot_longer(immuno:others, names_to = NULL, values_to = 'type', values_drop_na = T) %>%
  rename(NCTNumber = `NCT Number`) %>%
  plyr::ddply('NCTNumber', function(dfsub){
    types <- dfsub$type %>% unique() %>% sort() %>% str_c(collapse = ';')
    dfsub %>% slice(1) %>% mutate(type = all_of(types))
  })
  
df_drug_types %>% count(type)
df_drug_types %>%
  plyr::dlply('type', function(dfsub){
    dfsub$drug %>% unique() %>% sort()
  })
df_cli$drugtype <- df_drug_types$type



df_drug_prot_wide <- plyr::ddply(df_drug_prot, 'drug', function(dfsub){
  dfret <- dfsub %>% slice(1)
  dfret$UniprotID <- dfsub$UniprotID %>% unique() %>% sort() %>% str_c(collapse = ';')
  return(dfret)
})

ht_drug2prot <- df_drug_prot$UniprotID
names(ht_drug2prot) <- df_drug_prot$drug

df_cli$target <- sapply(df_cli$drug, function(e){
  drugs <- str_split(e, ';')[[1]]
  ht_drug2prot[drugs] %>%
    str_split(';') %>%
    unlist() %>%
    unique() %>%
    sort() %>%
    str_c(collapse = ';')
})



df_cli %>% rio::export('//172.16.13.136/share/members/jiangwenhao/TPHP/20220908/clinical_trials_download/TPHP_beast_endometrial_cancer_clinical_trials_hasResult_Interventional_phaseApplicable_without_nuterentional_antitumorDrug_748trials_drugType.tsv')



df_cli_1_2 <- readxl::read_excel('//172.16.13.136/share/members/jiangwenhao/TPHP/20220908/TPHP_carcinoma_upregulated_overlapped_drugbank/TPHP_carcinoma_upregulated_proteins_drugbank_clinical_trials_3754_cancertumor_breastEndometrial_combBothOne_20221110.xlsx')
df_cli_1_2_simple <- df_cli_1_2 %>% select(Name, Description, `Target name`, `Target actions`, `Target uniprot`)
intersect(colnames(df_cli_1_2_simple), colnames(df_cli)) # character(0)
df_cli_1_2_simple %<>%
  rename(drug = Name,
         DrugDescription = Description) %>%
  distinct()

ht_drug2target <- df_cli_1_2_simple$`Target uniprot`
names(ht_drug2target) <- df_cli_1_2_simple$drug

df_cli$`Target uniprot` <- sapply(df_cli$drug, function(e){
  drugs <- str_split(e, ';')[[1]]
  ht_drug2target[drugs] %>%
    str_split(';') %>%
    unlist() %>%
    unique() %>%
    str_trim() %>%
    .[. != ''] %>%
    sort() %>%
    str_c(collapse = ';')
})
df_cli
which(df_cli$target == df_cli$`Target uniprot`)



df_horm1 <- readxl::read_excel('//172.16.13.136/share/members/jiangwenhao/TPHP/20220908/clinical_trials_download/drugbank_results_hormoral.xlsx', sheet = 'drugbank_results_hormoral')
df_horm1_bio <- df_horm1 %>% filter(Biotech, !`Small molecule`, !Nutraceutical)


horm1_bio <- df_horm1_bio$Name
df_cli$IsBiotech <- sapply(df_cli$drug, function(e){
  str_split(e, ';')[[1]] %>%
    intersect(horm1_bio) %>%
    unique() %>% sort() %>% str_c(collapse = ';')
}) %>% unname()
df_cli$IsBiotech %>% .[. != ''] %>%
  str_split(';') %>%
  unlist() %>%
  table()
df_cli %>% rio::export('//172.16.13.136/share/members/jiangwenhao/TPHP/20220908/clinical_trials_download/TPHP_beast_endometrial_cancer_clinical_trials_hasResult_Interventional_phaseApplicable_without_nuterentional_antitumorDrug_748trials_drugTypeTarget.tsv')

drugs <- df_cli$IsBiotech %>% .[. != ''] %>%
  str_split(';') %>%
  unlist() %>%
  unique() %>% sort()

df_biodrug <- df_cli_1_2 %>% filter(Name %in% drugs) %>% select(Name, UniprotID:`Target uniprot`) %>% distinct()
df_biodrug %>%
  rio::export('//172.16.13.136/share/members/jiangwenhao/TPHP/20220908/clinical_trials_download/TPHP_beast_endometrial_cancer_clinical_trials_hasResult_Interventional_phaseApplicable_without_nuterentional_antitumorDrug_748trials_BiotechDrugs.xlsx')


# df_cli_simple <- df_cli %>% select(`NCT Number`, Title, Status, Conditions, Interventions, Phases, `Start Date`:`Last Update Posted`, cancer:target)
# df_cli_simple %>% 
#   rio::export('//172.16.13.136/share/members/jiangwenhao/TPHP/20220908/clinical_trials_download/TPHP_beast_endometrial_cancer_clinical_trials_hasResult_Interventional_phaseApplicable_without_nuterentional_antitumorDrug_748trials_drugType_simple.tsv')


df_cli_simple <- df_cli %>% select(`NCT Number`, Title, Status, Conditions, Interventions, Phases, `Start Date`:`Last Update Posted`, cancer:IsBiotech)
df_cli_simple %>% 
  rio::export('//172.16.13.136/share/members/jiangwenhao/TPHP/20220908/clinical_trials_download/TPHP_beast_endometrial_cancer_clinical_trials_hasResult_Interventional_phaseApplicable_without_nuterentional_antitumorDrug_748trials_BiotechDrugs_simple.tsv')



df_cli_simple %>% count(Phases, drugtype) %>%
  pivot_wider(Phases, names_from = drugtype, values_from = n)


df_drug_label %>% filter(others) %>% pull(drug)
# 'Ceralasertib;MK-1775;Olaparib'
df_target_uniprot_drug %>% filter(drug %in% c('Ceralasertib', 'MK-1775', 'Olaparib')) %>% pull(drug) %>% unique()
# without Ceralasertib



# clinicals with targets among hormoral, immuno and rtk
df_hir <- df_cli_simple %>%
  filter(drugtype == 'hormoral;immuno;rtk')

df_hir %>% 
  rio::export('//172.16.13.136/share/members/jiangwenhao/TPHP/20220908/clinical_trials_download/TPHP_beast_endometrial_cancer_clinical_trials_hasResult_Interventional_phaseApplicable_without_nuterentional_antitumorDrug_748trials_drugType_simple_hormoral-immuno-rtk.tsv')

df_hir %>% count(Phases, cancer) %>% pivot_wider(Phases, names_from = cancer, values_from = n)



# different targets between endometrial and breast cancers among hormoral, immuno and rtk
prot_hir_bre <- df_hir %>% 
  filter(cancer == 'Breast Cancer') %>%
  pull(target) %>%
  str_split(';') %>%
  unlist() %>% unique()

prot_hir_end <- df_hir %>% 
  filter(cancer == 'Endometrial Cancer') %>%
  pull(target) %>%
  str_split(';') %>%
  unlist() %>% unique()

setdiff(prot_hir_bre, prot_hir_end)
setdiff(prot_hir_end, prot_hir_bre)



# different targets between endometrial and breast cancers
drug_bre <- df_cli %>% 
  filter(cancer == 'Breast Cancer') %>%
  pull(drug) %>%
  str_split(';') %>%
  unlist() %>% unique()

drug_end <- df_cli %>% 
  filter(cancer == 'Endometrial Cancer') %>%
  pull(drug) %>%
  str_split(';') %>%
  unlist() %>% unique()

setdiff(drug_bre, drug_end) %>% sort()
setdiff(drug_end, drug_bre) %>% sort()
intersect(drug_bre, drug_end) %>% sort()





```









## pancreas cancer
### read drug pairs
```{r}
df_drug_pair <- readxl::read_excel('//172.16.13.136/share/members/jiangwenhao/TPHP/20220908/clinical_trials_download/pancreas/Final_pancreas_list_v2.xlsx', sheet = 'drug_combination')
drugs <- apply(df_drug_pair, 1, function(row){
  row %>% sort() %>% str_c(collapse = '&')
}) %>% unique()

```


### crawl clinical trials
```{python}
import requests
import re
import os
import pandas as pd
from tqdm import tqdm


drugs = r.drugs
cancers = ['Pancreas+Cancer']

rlt_dir = '//172.16.13.136/share/members/jiangwenhao/TPHP/20220908/clinical_trials_download/pancreas/tables/'
if not os.path.exists(rlt_dir): # output folder
    os.mkdir(rlt_dir)
for cancer in cancers:
    for drug in tqdm(drugs):
        # url = f'https://clinicaltrials.gov/ct2/results?type=&rslt=With&cond={cancer}&term={drug}'
        tb = f'https://clinicaltrials.gov/ct2/results/download_fields?down_count=10000&down_flds=all&down_fmt=csv&term={drug}&rslt=With&cond={cancer}&flds=a&flds=b&flds=e&flds=y'
        response = requests.get(tb)
        content = response.content.decode(response.encoding).encode('utf-8')
        with open(f'{rlt_dir}/{cancer}_{drug}.csv', 'wb') as f:
            f.write(content)


```



### combine tables
```{r}
tbls_cli <- list.files('//172.16.13.136/share/members/jiangwenhao/TPHP/20220908/clinical_trials_download/pancreas/tables/', full.names = T)
tbl_ls <- lapply(tbls_cli, function(fname){
  cancer <- fname %>% str_split('/') %>% unlist() %>% tail(1) %>%
    str_extract('.+?_') %>%
    str_replace('_', '') %>%
    str_replace('\\+', ' ')
  drug <- fname %>% str_split('/') %>% unlist() %>% tail(1) %>%
    str_extract('_.+?\\.csv') %>%
    str_replace('_', '') %>%
    str_replace('\\.csv', '')
  dftmp <- read.csv(fname, check.names = F)
  dftmp %<>% mutate(
    cancer = all_of(cancer),
    drug = all_of(drug)
  )
  return(dftmp)
})
df_cli <- plyr::ldply(tbl_ls)
df_cli %<>% select(`NCT Number`:drug)
df_cli %<>% filter(str_detect(`NCT Number`, '^NCT')) # remove unused rows

df_cli %>% rio::export('//172.16.13.136/share/members/jiangwenhao/TPHP/20220908/clinical_trials_download/pancreas/TPHP_pancreas_cancer_combinedDrugs_clinical_trials_hasResult.tsv')
df_cli <- rio::import('//172.16.13.136/share/members/jiangwenhao/TPHP/20220908/clinical_trials_download/pancreas/TPHP_pancreas_cancer_combinedDrugs_clinical_trials_hasResult.tsv')
df_cli$`NCT Number` %>% unique %>% length # 25

# nutrientional <- rio::import('//172.16.13.136/share/members/jiangwenhao/TPHP/20220908/clinical_trials_download/nutrientional_drugbank_results.csv') %>% pull(Name)
# df_cli %<>% filter(!(drug %in% nutrientional))
# df_cli %<>% filter(`Study Type` == 'Interventional')
# df_cli %<>% filter(Phases != 'Not Applicable')
# df_cli %>% rio::export('//172.16.13.136/share/members/jiangwenhao/TPHP/20220908/clinical_trials_download/TPHP_beast_endometrial_cancer_clinical_trials_hasResult_Interventional_phaseApplicable_without_nuterentional.tsv')
# 

```


## pancreas cancer with has no results
### read drug pairs
```{r}
df_drug_pair <- readxl::read_excel('//172.16.13.136/share/members/jiangwenhao/TPHP/20220908/clinical_trials_download/pancreas/Final_pancreas_list_v2.xlsx', sheet = 'drug_combination')
drugs <- apply(df_drug_pair, 1, function(row){
  row %>% sort() %>% str_c(collapse = '&')
}) %>% unique()

```

### crawl clinical trials
```{python}
import requests
import re
import os
import pandas as pd
from tqdm import tqdm


drugs = r.drugs
cancers = ['Pancreas+Cancer']

rlt_dir = '//172.16.13.136/share/members/jiangwenhao/TPHP/20220908/clinical_trials_download/pancreas/tables_WithOrWithout_results/'
if not os.path.exists(rlt_dir): # output folder
    os.mkdir(rlt_dir)
for cancer in cancers:
    for drug in tqdm(drugs):
        # url = f'https://clinicaltrials.gov/ct2/results?type=&rslt=With&cond={cancer}&term={drug}'
        tb = f'https://clinicaltrials.gov/ct2/results/download_fields?down_count=10000&down_flds=all&down_fmt=csv&term={drug}&cond={cancer}&flds=a&flds=b&flds=e&flds=y'
        response = requests.get(tb)
        content = response.content.decode(response.encoding).encode('utf-8')
        with open(f'{rlt_dir}/{cancer}_{drug}.csv', 'wb') as f:
            f.write(content)


```



### combine tables
```{r}
tbls_cli <- list.files('//172.16.13.136/share/members/jiangwenhao/TPHP/20220908/clinical_trials_download/pancreas/tables_WithOrWithout_results/', full.names = T)
tbl_ls <- lapply(tbls_cli, function(fname){
  cancer <- fname %>% str_split('/') %>% unlist() %>% tail(1) %>%
    str_extract('.+?_') %>%
    str_replace('_', '') %>%
    str_replace('\\+', ' ')
  drug <- fname %>% str_split('/') %>% unlist() %>% tail(1) %>%
    str_extract('_.+?\\.csv') %>%
    str_replace('_', '') %>%
    str_replace('\\.csv', '')
  dftmp <- read.csv(fname, check.names = F)
  dftmp %<>% mutate(
    cancer = all_of(cancer),
    drug = all_of(drug)
  )
  return(dftmp)
})
df_cli <- plyr::ldply(tbl_ls)
df_cli %<>% select(`NCT Number`:drug)
df_cli %<>% filter(str_detect(`NCT Number`, '^NCT')) # remove unused rows

df_cli %>% rio::export('//172.16.13.136/share/members/jiangwenhao/TPHP/20220908/clinical_trials_download/pancreas/TPHP_pancreas_cancer_combinedDrugs_clinical_trials_WithOrWithout_results.tsv')
df_cli <- rio::import('//172.16.13.136/share/members/jiangwenhao/TPHP/20220908/clinical_trials_download/pancreas/TPHP_pancreas_cancer_combinedDrugs_clinical_trials_WithOrWithout_results.tsv')
df_cli$`NCT Number` %>% unique %>% length # 190

# nutrientional <- rio::import('//172.16.13.136/share/members/jiangwenhao/TPHP/20220908/clinical_trials_download/nutrientional_drugbank_results.csv') %>% pull(Name)
# df_cli %<>% filter(!(drug %in% nutrientional))
# df_cli %<>% filter(`Study Type` == 'Interventional')
# df_cli %<>% filter(Phases != 'Not Applicable')
# df_cli %>% rio::export('//172.16.13.136/share/members/jiangwenhao/TPHP/20220908/clinical_trials_download/TPHP_beast_endometrial_cancer_clinical_trials_hasResult_Interventional_phaseApplicable_without_nuterentional.tsv')
# 

```






# 2. GDSC resource
## data preparation
```{r}
df_dose1 <- rio::import('//172.16.13.136/share/members/jiangwenhao/TPHP/20220908/GDSC_resource/release-8.4/GDSC1_fitted_dose_response_24Jul22.csv')
df_dose2 <- rio::import('//172.16.13.136/share/members/jiangwenhao/TPHP/20220908/GDSC_resource/release-8.4/GDSC2_fitted_dose_response_24Jul22.csv')

# identical(colnames(df_dose1), colnames(df_dose2)) # TRUE
df_dose <- rbind(df_dose1, df_dose2)


# combine with cell_line table
df_cosmic <-  readxl::read_excel('//172.16.13.136/share/members/jiangwenhao/TPHP/20220908/GDSC_resource/release-8.4/Cell_Lines_Details.xlsx', sheet = 'COSMIC tissue classification')
df_merge1 <- full_join(df_cosmic, df_dose, by = 'COSMIC_ID')


# map gene to UniprotID 
df_gene <- df_merge1 %>% select(PUTATIVE_TARGET) %>% distinct()
UniprotID <- c()
for(i in 1:length(df_gene$PUTATIVE_TARGET)){
  print(str_glue('Search gene {i}/{length(df_gene$PUTATIVE_TARGET)}'))
  gene_symbol <- df_gene$PUTATIVE_TARGET[i]
  gene_symbols <- str_split(gene_symbol, ', ')[[1]]
  uniprot_id <- sapply(gene_symbols, genesymbol_to_uniprotid)
  UniprotID[i] <- str_c(uniprot_id, collapse = ', ')
}
df_gene$UniprotID <- UniprotID
rio::export(df_gene, '//172.16.13.136/share/members/jiangwenhao/TPHP/20220908/TPHP_carcinoma_upregulated_overlapped_drugbank/GDSC_target_to_uniprotid.xlsx')


# combine with GeneSymbol_UniprotID table
df_gene <- rio::import('//172.16.13.136/share/members/jiangwenhao/TPHP/20220908/TPHP_carcinoma_upregulated_overlapped_drugbank/GDSC_target_to_uniprotid_edited.xlsx')
df_merge2 <- full_join(df_merge1, df_gene, by = 'PUTATIVE_TARGET') %>%
  relocate(UniprotID, .after = PUTATIVE_TARGET)
rio::export(df_merge2, '//172.16.13.136/share/members/jiangwenhao/TPHP/20220908/TPHP_carcinoma_upregulated_overlapped_drugbank/GDSC_fitted_dose_response_add_uniprotid.xlsx')


# filter with upregulated proteins
file_ls <- list.files('//172.16.13.136/share/members/yuel/2022/dysregulated proteins/',
                      pattern = '*\\.xlsx$', full.names = T)
organs <- str_extract(file_ls, '05p_.+\\.xlsx') %>%
  str_replace('05p_', '') %>% str_replace('\\.xlsx', '')

tbl1_ls <- lapply(file_ls, readxl::read_excel, sheet = 1)
tbl_ls <- lapply(seq_along(file_ls), function(i){# up only
  df_up <- tbl1_ls[[i]]
})
names(tbl_ls) <- organs
df_up <- plyr::ldply(tbl_ls, function(dfsub){
  dfsub
}, .id = 'organ')
prot_up <- unique(df_up$prot)

df_gene_up <- 
  df_gene %>%
  filter(
    str_detect(UniprotID, pattern = str_c(prot_up, collapse = '|'))
  )
df_gene_up$UniprotID <- str_split(df_gene_up$UniprotID, '(,)|(, )|(;)') %>%
  lapply(function(x){
    str_trim(x) %>% .[. != '-'] %>% unique() %>% sort() %>% str_c(collapse = ';')
  }) %>%
  unlist()
df_merge2up <- inner_join(df_merge1, df_gene_up, by = 'PUTATIVE_TARGET') %>%
  relocate(UniprotID, .after = PUTATIVE_TARGET)
rio::export(df_merge2up, '//172.16.13.136/share/members/jiangwenhao/TPHP/20220908/TPHP_carcinoma_upregulated_overlapped_drugbank/GDSC_fitted_dose_response_tphp-up_add_uniprotid.xlsx')

# df_gene_up %>% separate_rows('UniprotID') %>%
#   distinct() %>%
#   group_by(UniprotID) %>%
#   summarise(PUTATIVE_TARGET)

```


## ANOVA tables
```{r}
ano1_ls <- read_excel_allsheets('//172.16.13.136/share/members/jiangwenhao/TPHP/20220908/GDSC_resource/release-8.4/ANOVA_results_GDSC1_24Jul22.xlsx')
types <- ano1_ls[[1]] %>% # ANOVA tables2 required assigned column types
  summarise_all(typeof) %>%
  unlist() %>% unname() %>%
  str_replace_all('double', 'numeric') %>%
  str_replace_all('character', 'text')
ano2_ls <- read_excel_allsheets('//172.16.13.136/share/members/jiangwenhao/TPHP/20220908/GDSC_resource/release-8.4/ANOVA_results_GDSC2_24Jul22.xlsx', col_types = types)


df_ano1 <- plyr::ldply(ano1_ls)
df_ano2 <- plyr::ldply(ano2_ls)
df_ano <- rbind(df_ano1, df_ano2) %>% select(-.id)

df_merge2up <- rio::import('//172.16.13.136/share/members/jiangwenhao/TPHP/20220908/TPHP_carcinoma_upregulated_overlapped_drugbank/GDSC_fitted_dose_response_tphp-up_add_uniprotid.xlsx')
df_ano_up <- df_merge2up %>%
  select(PUTATIVE_TARGET, UniprotID) %>%
  rename(target = PUTATIVE_TARGET) %>%
  distinct() %>%
  inner_join(df_ano, by = 'target')

# tissue label
df_label <- readxl::read_excel('//172.16.13.136/share/members/jiangwenhao/TPHP/20220908/GDSC_resource/release-8.4/Cell_Lines_Details.xlsx', sheet = 'Decode')
df_label %<>% slice(2:31) %>% setNames(c('tissue_type', 'TCGA Definition'))
df_ano_up$tissue_type %>% unique() %>% sort() %>% setdiff(df_label$tissue_type)
df_label %<>% add_row(tissue_type = 'PANCANCER', `TCGA Definition` = 'PANCANCER')
rio::export(df_label, '//172.16.13.136/share/members/jiangwenhao/TPHP/20220908/TPHP_carcinoma_upregulated_overlapped_drugbank/GDSC_tissue_label.xlsx')

df_label <- rio::import('//172.16.13.136/share/members/jiangwenhao/TPHP/20220908/TPHP_carcinoma_upregulated_overlapped_drugbank/GDSC_tissue_label_edited.xlsx')
df_ano_up %<>% inner_join(df_label, ., by = 'tissue_type')

rio::export(df_ano_up, '//172.16.13.136/share/members/jiangwenhao/TPHP/20220908/TPHP_carcinoma_upregulated_overlapped_drugbank/GDSC_ANOVA_tphp-up.xlsx')


# ANOVA filter
df_ano_up_fdr <- df_ano_up %>% filter(fdr < 0.01)
rio::export(df_ano_up_fdr, '//172.16.13.136/share/members/jiangwenhao/TPHP/20220908/TPHP_carcinoma_upregulated_overlapped_drugbank/GDSC_ANOVA_tphp-up_fdr.xlsx')


df_ano_up_fdr1 <- df_ano_up_fdr %>%
  select(`TCGA Definition`, GDSC_organ, UniprotID, drug_name) %>% distinct() %>%
  separate_rows('UniprotID') %>%
  filter(UniprotID %in% prot_up) %>%
  distinct() %>%
  rename(prot = UniprotID) %>%
  inner_join(df_up[, 1:2], by = 'prot')
rio::export(df_ano_up_fdr1, '//172.16.13.136/share/members/jiangwenhao/TPHP/20220908/TPHP_carcinoma_upregulated_overlapped_drugbank/GDSC_ANOVA_tphp-up_fdr_simple.xlsx')

df_ano_up_fdr2 <- df_ano_up_fdr1 %>% filter(str_detect(GDSC_organ, as.character(organ)))
rio::export(df_ano_up_fdr2, '//172.16.13.136/share/members/jiangwenhao/TPHP/20220908/TPHP_carcinoma_upregulated_overlapped_drugbank/GDSC_ANOVA_tphp-up_fdr_simple_overlap.xlsx')



# definitions <- df_ano_up_fdr1$`TCGA Definition` %>% unique() %>% sort()
ano_up_fdr1_ls <- df_ano_up_fdr1 %>% select(`TCGA Definition`, prot) %>% distinct() %>%
  setNames(c('TCGA', 'prot')) %>%
  plyr::dlply(.variables = 'TCGA', function(dfsub){
    dfsub$prot
  })
UpSetR::upset(UpSetR::fromList(ano_up_fdr1_ls), order.by = "freq")
```





# new --
## carcinoma upregulated proteins overlap
```{r}
file_ls <- list.files('//172.16.13.136/share/members/yuel/2022/dysregulated proteins/',
                      pattern = '*\\.xlsx$', full.names = T)
organs <- str_extract(file_ls, '05p_.+\\.xlsx') %>%
  str_replace('05p_', '') %>% str_replace('\\.xlsx', '')

tbl1_ls <- lapply(file_ls, readxl::read_excel, sheet = 1)
tbl_ls <- lapply(seq_along(file_ls), function(i){# up only
  df_up <- tbl1_ls[[i]]
})
names(tbl_ls) <- organs
df_up <- plyr::ldply(tbl_ls, function(dfsub){
  dfsub
}, .id = 'organ')
prots_up <- sort(unique(df_up$prot))

df_up1 <- df_up %>% select(organ, prot)
up1_ls <- plyr::dlply(df_up1, 'organ', function(dfsub){
  dfsub$prot
})
UpSetR::upset(UpSetR::fromList(up1_ls), order.by = "freq", nsets = 2)

df_up_ <- df_up %>% filter(organ %in% c('mammary gland', 'uterus'))

```





## ANOVA tables
```{r}
ano1_ls <- read_excel_allsheets('//172.16.13.136/share/members/jiangwenhao/TPHP/20220908/GDSC_resource/release-8.4/ANOVA_results_GDSC1_24Jul22.xlsx')
types <- ano1_ls[[1]] %>% # ANOVA tables2 required assigned column types
  summarise_all(typeof) %>%
  unlist() %>% unname() %>%
  str_replace_all('double', 'numeric') %>%
  str_replace_all('character', 'text')
ano2_ls <- read_excel_allsheets('//172.16.13.136/share/members/jiangwenhao/TPHP/20220908/GDSC_resource/release-8.4/ANOVA_results_GDSC2_24Jul22.xlsx', col_types = types)


df_ano1 <- plyr::ldply(ano1_ls)
df_ano2 <- plyr::ldply(ano2_ls)
df_ano <- rbind(df_ano1, df_ano2) %>% select(-.id)
df_ano_gl <- df_ano %>% filter(str_detect(feature_name, '(gain\\.)|(loss\\.)')) # with gain/loss in feature




df_merge2up <- rio::import('//172.16.13.136/share/members/jiangwenhao/TPHP/20220908/TPHP_carcinoma_upregulated_overlapped_drugbank/GDSC_fitted_dose_response_tphp-up_add_uniprotid.xlsx')
df_ano_up <- df_merge2up %>%
  select(PUTATIVE_TARGET, UniprotID) %>%
  rename(target = PUTATIVE_TARGET) %>%
  distinct() %>%
  inner_join(df_ano_gl, by = 'target')

# # tissue label
# df_label <- readxl::read_excel('//172.16.13.136/share/members/jiangwenhao/TPHP/20220908/GDSC_resource/release-8.4/Cell_Lines_Details.xlsx', sheet = 'Decode')
# df_label %<>% slice(2:31) %>% setNames(c('tissue_type', 'TCGA Definition'))
# df_ano_up$tissue_type %>% unique() %>% sort() %>% setdiff(df_label$tissue_type)
# df_label %<>% add_row(tissue_type = 'PANCANCER', `TCGA Definition` = 'PANCANCER')
# rio::export(df_label, '//172.16.13.136/share/members/jiangwenhao/TPHP/20220908/TPHP_carcinoma_upregulated_overlapped_drugbank/GDSC_tissue_label.xlsx')

df_label <- rio::import('//172.16.13.136/share/members/jiangwenhao/TPHP/20220908/TPHP_carcinoma_upregulated_overlapped_drugbank/GDSC_tissue_label_edited.xlsx') %>% select(1:3)
df_ano_up %<>% inner_join(df_label, ., by = 'tissue_type')
# prots_up_ptn <- str_c(prots_up, collapse = '|')
df_ano_up %<>% filter(`TCGA Definition` != 'PANCANCER', # without PANCANCER
                      # str_detect(UniprotID, prots_up_ptn)
)
rio::export(df_ano_up, '//172.16.13.136/share/members/jiangwenhao/TPHP/20220908/TPHP_carcinoma_upregulated_overlapped_drugbank/GDSC_ANOVA_LOSSorGAIN_tphp-up.xlsx')


df_ano_up_sim <- df_ano_up %>%
  select(drug_name, tissue_type, `TCGA Definition`, GDSC_organ, target, UniprotID, feature_name, feature_delta_mean_ic50)
rio::export(df_ano_up_sim , '//172.16.13.136/share/members/jiangwenhao/TPHP/20220908/TPHP_carcinoma_upregulated_overlapped_drugbank/GDSC_ANOVA_LOSSorGAIN_tphp-up_simple.xlsx')







# 比较相同 target 情况下，不同 organ 的 feature_delta_mean_ic50含量差异，及其在 TPHP data 中是否为相应 organ 的 upregulated protein
df_ano_up_sim %<>% arrange(target, tissue_type) %>%
  filter(!is.na(GDSC_organ))



```



<!-- # ```{r} -->
<!-- # df_ano_up_sim_test <- df_ano_up_sim %>% filter(target == 'ABL') -->
<!-- # df_ano_up_sim_test$tissue_type %<>% factor() -->
<!-- #  -->
<!-- # # df_up_ %>% filter(prot == 'P00519') -->
<!-- # logmean <- function(x){log(mean(exp(x)))} -->
<!-- #  -->
<!-- # df_ano_up_sim_test %>% -->
<!-- #   group_by(tissue_type, GDSC_organ) %>% -->
<!-- #   summarise(mean = logmean(feature_delta_mean_ic50)) -->
<!-- #  -->
<!-- # df_ano_up_sim_test %>% -->
<!-- #   group_by(tissue_type, GDSC_organ, feature_name) %>% -->
<!-- #   summarise(mean = logmean(feature_delta_mean_ic50)) -->
<!-- #  -->
<!-- #  -->
<!-- #  -->
<!-- #  -->
<!-- #  -->
<!-- # # line only -->
<!-- # df_plot <- df_ano_up_sim_test %>% -->
<!-- #   group_by(tissue_type) %>% -->
<!-- #   summarise(mean = logmean(feature_delta_mean_ic50)) %>% -->
<!-- #   arrange(desc(mean)) %>% -->
<!-- #   mutate(tissue_type = factor(tissue_type, levels = unique(.$tissue_type))) %>% -->
<!-- #   mutate(rank = 1:nrow(.)) -->
<!-- #  -->
<!-- # p1 <- ggplot(df_plot, aes(x = tissue_type, y = mean)) + -->
<!-- #   # geom_line(aes(tissue_type)) + -->
<!-- #   geom_point(aes(color = tissue_type)) + -->
<!-- #   geom_smooth(mapping = aes(x = rank, y = mean), data = df_plot, method = 'loess', color = '#487DF0') + -->
<!-- #   theme_bw()+ -->
<!-- #     theme(legend.text = element_text(size = 18, color = "black"), -->
<!-- #           legend.position = 'top', -->
<!-- #           legend.title = element_text(size = 18, color="black") , -->
<!-- #           panel.grid.major =element_blank(), -->
<!-- #           panel.grid.minor = element_blank(), -->
<!-- #           panel.background = element_blank(), -->
<!-- #           axis.line = element_line(colour = "black"))+ -->
<!-- #     theme(panel.grid =element_blank())+ -->
<!-- #     theme(axis.text = element_text(size = 15,color = "black"))+ -->
<!-- #     theme(axis.text.x = element_text(size = 15,color = "black", angle = 0))+ -->
<!-- #     theme(plot.title = element_text(size = 20, face = "bold"))+ -->
<!-- #     theme(plot.subtitle=element_text(size = 20, hjust = 0, color="black"))+ -->
<!-- #     theme(axis.title.x=element_text(size = 18, hjust = 0.5, color="black"))+ -->
<!-- #     theme(axis.title.y=element_text(size = 18, hjust = 0.5, color="black")) -->
<!-- #  -->
<!-- #  -->
<!-- #  -->
<!-- #  -->
<!-- # df_ano_up_sim_test %<>% -->
<!-- #   mutate(tissue_type = factor(tissue_type, levels = levels(df_plot$tissue_type))) %>% -->
<!-- #   arrange(desc(tissue_type)) -->
<!-- #  -->
<!-- #  -->
<!-- #  -->
<!-- # # scatter -->
<!-- # p2 <- ggplot(df_ano_up_sim_test, aes(x = feature_name, y = feature_delta_mean_ic50, group = tissue_type)) + -->
<!-- #   # geom_line(aes(tissue_type)) + -->
<!-- #   geom_point(aes(color = tissue_type)) + -->
<!-- #   theme_bw()+ -->
<!-- #     theme(legend.text = element_text(size = 18, color = "black"), -->
<!-- #           legend.position = 'top', -->
<!-- #           legend.title = element_text(size = 18, color="black") , -->
<!-- #           panel.grid.major =element_blank(), -->
<!-- #           panel.grid.minor = element_blank(), -->
<!-- #           panel.background = element_blank(), -->
<!-- #           axis.line = element_line(colour = "black"))+ -->
<!-- #     theme(panel.grid =element_blank())+ -->
<!-- #     theme(axis.text = element_text(size = 15,color = "black"))+ -->
<!-- #     theme(axis.text.x = element_text(size = 15,color = "black", angle = 0))+ -->
<!-- #     theme(plot.title = element_text(size = 20, face = "bold"))+ -->
<!-- #     theme(plot.subtitle=element_text(size = 20, hjust = 0, color="black"))+ -->
<!-- #     theme(axis.title.x=element_text(size = 18, hjust = 0.5, color="black"))+ -->
<!-- #     theme(axis.title.y=element_text(size = 18, hjust = 0.5, color="black")) -->
<!-- #  -->
<!-- # ggsave('//172.16.13.136/share/members/jiangwenhao/TPHP/20220908/TPHP_carcinoma_upregulated_overlapped_drugbank/GDSC_ANOVA_LOSSorGAIN_tphp-up_simple_test.pdf', p, width = 16, height = 9) -->
<!-- #  -->
<!-- #  -->
<!-- #  -->
<!-- # # combine -->
<!-- # p <-  -->
<!-- #   ggplot() + -->
<!-- #   # geom_line(aes(feature_name)) + -->
<!-- #   geom_point(mapping = aes(x = tissue_type, y = feature_delta_mean_ic50, group = feature_name, color = feature_name), -->
<!-- #              data = df_ano_up_sim_test) + -->
<!-- #   geom_smooth(mapping = aes(x = rank, y = mean), data = df_plot, method = 'loess', color = '#487DF0')+ -->
<!-- #   theme_bw()+ -->
<!-- #     theme(legend.text = element_text(size = 18, color = "black"), -->
<!-- #           legend.position = 'none', -->
<!-- #           legend.title = element_text(size = 18, color="black") , -->
<!-- #           panel.grid.major =element_blank(), -->
<!-- #           panel.grid.minor = element_blank(), -->
<!-- #           panel.background = element_blank(), -->
<!-- #           axis.line = element_line(colour = "black"))+ -->
<!-- #     theme(panel.grid =element_blank())+ -->
<!-- #     theme(axis.text = element_text(size = 15,color = "black"))+ -->
<!-- #     theme(axis.text.x = element_text(size = 15,color = "black", angle = 0))+ -->
<!-- #     theme(plot.title = element_text(size = 20, face = "bold"))+ -->
<!-- #     theme(plot.subtitle=element_text(size = 20, hjust = 0, color="black"))+ -->
<!-- #     theme(axis.title.x=element_text(size = 18, hjust = 0.5, color="black"))+ -->
<!-- #     theme(axis.title.y=element_text(size = 18, hjust = 0.5, color="black")) -->
<!-- #  -->
<!-- #  -->
<!-- # ``` -->


## map dysregulated protein to organs
```{r}
df_prot2organ <- df_up %>% select(organ, prot) %>%
  group_by(prot) %>%
  summarise(organ = str_c(sort(unique(organ)), collapse = ';'))

ano_prots <- lapply(df_ano_up_sim$UniprotID, function(x) str_split(x, ';')[[1]]) %>%
  unlist() %>% unique()
df_prot2organ1 <- df_prot2organ %>% filter(prot %in% ano_prots)
ht_prot2organ1 <- df_prot2organ1$organ
names(ht_prot2organ1) <- df_prot2organ1$prot
df_ano_up_sim$TPHP_organ <- sapply(df_ano_up_sim$UniprotID, function(x){
  xx <- str_split(x, ';')[[1]]
  ht_prot2organ1[xx] %>% na.omit() %>% str_c(collapse = ';') %>% str_split(';') %>% .[[1]] %>%
    unique() %>% sort() %>% str_c(collapse = ';')
})




drugs <- sort(unique(df_ano_up_sim$drug_name))
pdf('//172.16.13.136/share/members/jiangwenhao/TPHP/20220908/TPHP_carcinoma_upregulated_overlapped_drugbank/GDSC_ANOVA_LOSSorGAIN_overlapped_tphp-up_target_drugs.pdf', width = 16, height = 9)
for(i in seq_along(drugs)){
  drug <- drugs[i]
  df_this_drug <- df_ano_up_sim %>% filter(drug_name == all_of(drug))
  df_this_drug$tissue_type %<>% factor()
  
  logmean <- function(x){log(mean(exp(x)))} # function: mean of log
  
  organ_tphp <- df_this_drug$TPHP_organ[1]
  this_prot <- df_this_drug$UniprotID[1]
  this_target <- df_this_drug$target[1]
  this_title <- str_c(drug, organ_tphp, this_prot, this_target, sep = ' ~ ')
  
  
  tissue_types <- df_this_drug %>%
    filter(str_detect(df_this_drug$GDSC_organ, str_replace_all(organ_tphp, ';', '|'))) %>%
    pull(tissue_type) %>%
    unique()
  
  
  df_plot <- df_this_drug %>%
    group_by(tissue_type) %>%
    summarise(mean = logmean(feature_delta_mean_ic50)) %>%
    arrange(desc(mean)) %>%
    mutate(tissue_type = factor(tissue_type, levels = unique(.$tissue_type))) %>%
    mutate(rank = 1:nrow(.))
  
  
  
  df_this_drug %<>%
    mutate(tissue_type = factor(tissue_type, levels = levels(df_plot$tissue_type))) %>%
    arrange(desc(tissue_type)) %>%
    mutate(alpha = ifelse(.$tissue_type %in% all_of(tissue_types), 1, 0.5))
  
  p <- 
    ggplot() +
    # geom_line(aes(feature_name)) +
    geom_point(mapping = aes(x = tissue_type, y = feature_delta_mean_ic50, group = feature_name, color = feature_name, alpha = alpha),
               data = df_this_drug) +
    geom_smooth(mapping = aes(x = rank, y = mean), data = df_plot, method = 'loess', color = '#487DF0')+
    labs(subtitle = this_title,
         y = 'Feature delta mean IC50')+
    theme_bw()+
    theme(legend.text = element_text(size = 18, color = "black"),
          legend.position = 'none',
          legend.title = element_text(size = 18, color="black") ,
          panel.grid.major =element_blank(),
          panel.grid.minor = element_blank(),
          panel.background = element_blank(),
          axis.line = element_line(colour = "black"))+
    theme(panel.grid =element_blank())+
    theme(axis.text = element_text(size = 15,color = "black"))+
    theme(axis.text.x = element_text(size = 15,color = "black", angle = 90, vjust = 0.5))+
    theme(plot.title = element_text(size = 20, face = "bold"))+
    theme(plot.subtitle=element_text(size = 20, hjust = 0, color="black"))+
    theme(axis.title.x=element_blank())+
    theme(axis.title.y=element_text(size = 18, hjust = 0.5, color="black"))
  
  print(p)
  
}
graphics.off()

```


## gain & loss
```{r}
# LOSS
pdf('//172.16.13.136/share/members/jiangwenhao/TPHP/20220908/TPHP_carcinoma_upregulated_overlapped_drugbank/GDSC_ANOVA_LOSS_tphp-up_overlapped_target_drugs.pdf', width = 16, height = 9)
for(i in seq_along(drugs)){
  drug <- drugs[i]
  df_this_drug <- df_ano_up_sim %>%
    filter(str_detect(feature_name, 'loss\\.')) %>%
    filter(drug_name == all_of(drug))
  df_this_drug$tissue_type %<>% factor()
  
  logmean <- function(x){log(mean(exp(x)))} # function: mean of log
  
  organ_tphp <- df_this_drug$TPHP_organ[1]
  this_prot <- df_this_drug$UniprotID[1]
  this_target <- df_this_drug$target[1]
  this_title <- str_c(drug, organ_tphp, this_prot, this_target, sep = ' ~ ')
  
  
  tissue_types <- df_this_drug %>%
    filter(str_detect(df_this_drug$GDSC_organ, str_replace_all(organ_tphp, ';', '|'))) %>%
    pull(tissue_type) %>%
    unique()
  
  
  df_plot <- df_this_drug %>%
    group_by(tissue_type) %>%
    summarise(mean = logmean(feature_delta_mean_ic50)) %>%
    arrange(desc(mean)) %>%
    mutate(tissue_type = factor(tissue_type, levels = unique(.$tissue_type))) %>%
    mutate(rank = 1:nrow(.))
  
  
  
  df_this_drug %<>%
    mutate(tissue_type = factor(tissue_type, levels = levels(df_plot$tissue_type))) %>%
    arrange(desc(tissue_type)) %>%
    mutate(alpha = ifelse(.$tissue_type %in% all_of(tissue_types), 1, 0.5))
  
  p <- 
    ggplot() +
    # geom_line(aes(feature_name)) +
    geom_point(mapping = aes(x = tissue_type, y = feature_delta_mean_ic50, group = feature_name, color = feature_name, alpha = alpha),
               data = df_this_drug) +
    geom_smooth(mapping = aes(x = rank, y = mean), data = df_plot, method = 'loess', color = '#487DF0')+
    labs(subtitle = this_title,
         y = 'Feature delta mean IC50')+
    theme_bw()+
    theme(legend.text = element_text(size = 18, color = "black"),
          legend.position = 'none',
          legend.title = element_text(size = 18, color="black") ,
          panel.grid.major =element_blank(),
          panel.grid.minor = element_blank(),
          panel.background = element_blank(),
          axis.line = element_line(colour = "black"))+
    theme(panel.grid =element_blank())+
    theme(axis.text = element_text(size = 15,color = "black"))+
    theme(axis.text.x = element_text(size = 15,color = "black", angle = 90, vjust = 0.5))+
    theme(plot.title = element_text(size = 20, face = "bold"))+
    theme(plot.subtitle=element_text(size = 20, hjust = 0, color="black"))+
    theme(axis.title.x=element_blank())+
    theme(axis.title.y=element_text(size = 18, hjust = 0.5, color="black"))
  
  print(p)
  
}
graphics.off()




# GAIN
pdf('//172.16.13.136/share/members/jiangwenhao/TPHP/20220908/TPHP_carcinoma_upregulated_overlapped_drugbank/GDSC_ANOVA_GAIN_tphp-up_overlapped_target_drugs.pdf', width = 16, height = 9)
for(i in seq_along(drugs)){
  drug <- drugs[i]
  df_this_drug <- df_ano_up_sim %>%
    filter(str_detect(feature_name, 'gain\\.')) %>%
    filter(drug_name == all_of(drug))
  df_this_drug$tissue_type %<>% factor()
  
  logmean <- function(x){log(mean(exp(x)))} # function: mean of log
  
  organ_tphp <- df_this_drug$TPHP_organ[1]
  this_prot <- df_this_drug$UniprotID[1]
  this_target <- df_this_drug$target[1]
  this_title <- str_c(drug, organ_tphp, this_prot, this_target, sep = ' ~ ')
  
  
  tissue_types <- df_this_drug %>%
    filter(str_detect(df_this_drug$GDSC_organ, str_replace_all(organ_tphp, ';', '|'))) %>%
    pull(tissue_type) %>%
    unique()
  
  
  df_plot <- df_this_drug %>%
    group_by(tissue_type) %>%
    summarise(mean = logmean(feature_delta_mean_ic50)) %>%
    arrange(desc(mean)) %>%
    mutate(tissue_type = factor(tissue_type, levels = unique(.$tissue_type))) %>%
    mutate(rank = 1:nrow(.))
  
  
  
  df_this_drug %<>%
    mutate(tissue_type = factor(tissue_type, levels = levels(df_plot$tissue_type))) %>%
    arrange(desc(tissue_type)) %>%
    mutate(alpha = ifelse(.$tissue_type %in% all_of(tissue_types), 1, 0.5))
  
  p <- 
    ggplot() +
    # geom_line(aes(feature_name)) +
    geom_point(mapping = aes(x = tissue_type, y = feature_delta_mean_ic50, group = feature_name, color = feature_name, alpha = alpha),
               data = df_this_drug) +
    geom_smooth(mapping = aes(x = rank, y = mean), data = df_plot, method = 'loess', color = '#487DF0')+
    labs(subtitle = this_title,
         y = 'Feature delta mean IC50')+
    theme_bw()+
    theme(legend.text = element_text(size = 18, color = "black"),
          legend.position = 'none',
          legend.title = element_text(size = 18, color="black") ,
          panel.grid.major =element_blank(),
          panel.grid.minor = element_blank(),
          panel.background = element_blank(),
          axis.line = element_line(colour = "black"))+
    theme(panel.grid =element_blank())+
    theme(axis.text = element_text(size = 15,color = "black"))+
    theme(axis.text.x = element_text(size = 15,color = "black", angle = 90, vjust = 0.5))+
    theme(plot.title = element_text(size = 20, face = "bold"))+
    theme(plot.subtitle=element_text(size = 20, hjust = 0, color="black"))+
    theme(axis.title.x=element_blank())+
    theme(axis.title.y=element_text(size = 18, hjust = 0.5, color="black"))
  
  print(p)
  
}
graphics.off()


```


# 2. GDSC use S4C table
```{r}
dfs4c <- rio::import('//172.16.13.136/share/members/jiangwenhao/TPHP/20220908/TPHP_carcinoma_upregulated_overlapped_drugbank/TableS4C-ANOVAsignificantHits.xlsx')
dfs4c1 <- dfs4c %>% filter(str_detect(FEATURE, 'gain:')) # select gain
dfs4c2 <- dfs4c1 %>% filter(str_detect(FEATURE, '\\(.+?\\)')) # select which with target name


# map gene to UniprotID 
gene_symbols <- str_extract_all(dfs4c2$FEATURE, '\\(.+?\\)') %>%
  unlist() %>% unique() %>%
  str_replace_all('\\(|\\)', '') %>%
  str_c(collapse = ',') %>%
  str_split(',') %>% .[[1]] %>%
  unique()
uniprotids <- c()
for(i in seq_along(gene_symbols)){
  gene_symbol <- gene_symbols[i]
  print(str_glue('{gene_symbol}  {i}/{length(gene_symbols)}'))
  uniprotids[i] <- genesymbol_to_uniprotid(gene_symbol)
}

ht_gene2prot <- uniprotids
names(ht_gene2prot) <- gene_symbols


dfs4c3 <- dfs4c2 %>%
  add_column(
    FEATURE_GENE = sapply(dfs4c2$FEATURE, function(x){
      x %>%
        str_extract_all('\\(.+?\\)') %>%
        unlist() %>% unique() %>%
        str_replace_all('\\(|\\)', '') %>%
        str_c(collapse = ',') %>%
        str_split(',') %>% .[[1]] %>%
        unique() %>%
        str_c(collapse = ',')
    }),
    FEATURE_PROTEIN = sapply(dfs4c2$FEATURE, function(x){
      x %>%
        str_extract_all('\\(.+?\\)') %>%
        unlist() %>% unique() %>%
        str_replace_all('\\(|\\)', '') %>%
        str_c(collapse = ',') %>%
        str_split(',') %>% .[[1]] %>%
        unique() %>%
        ht_gene2prot[.] %>%
        str_c(collapse = ',')
    }),
    .before = 'FEATURE'
  )
dfs4c3 %>% rio::export('//172.16.13.136/share/members/jiangwenhao/TPHP/20220908/TPHP_carcinoma_upregulated_overlapped_drugbank/TableS4C-ANOVAsignificantHits_addcolumn_gene_protein.xlsx')

dfs4c3[dfs4c3$Domain == 'COAD/READ', 'Domain'] <- 'COREAD'
dfs4c4 <- dfs4c3 %>%
  rename(tissue_type = Domain) %>%
  right_join(df_label, ., by = 'tissue_type')
dfs4c4[dfs4c4$tissue_type == 'PANCAN', 'GDSC_organ'] <- 'PANCAN'

organ_prot_ls <- df_tbls %>% select(organ, UniprotID) %>% distinct() %>%
  plyr::dlply('organ', pull, UniprotID)
organ_prot_ls[['PANCAN']] <- unique(df_tbls$UniprotID)

dfs4c4 %<>% add_column(
  FEATURE_PROTEIN_mapped = apply(dfs4c4, 1, function(row){
    if(is.na(row['GDSC_organ'])){
      prots <- NA
    } else{
      prots <- row['GDSC_organ'] %>% str_split(';') %>% .[[1]] %>%
        lapply(function(e) organ_prot_ls[[e]]) %>%
        unlist() %>% unique()
    }
    feature_prots <- row['FEATURE_PROTEIN'] %>% str_split(',') %>% .[[1]]
    prot_inter <- intersect(prots, feature_prots)
    ret <- str_c(prot_inter, collapse = ';')
    ret <- ifelse(ret == '', NA, ret)
    return(ret)
  }),
  .after = 'FEATURE_PROTEIN'
)
prot_organ_ls <- df_tbls %>% select(organ, UniprotID) %>% distinct() %>%
  plyr::dlply('UniprotID', function(dfsub){
    dfsub %>% pull(organ)
  })


dfs4c4 %<>% add_column(
  FEATURE_PROTEIN_mapped_organ = sapply(dfs4c4$FEATURE_PROTEIN_mapped, function(x){
    if(is.na(x)){
      ret <- NA
    } else{
      ret <- str_split(x, ';') %>% unlist() %>% lapply(function(e) prot_organ_ls[[e]] %>% str_c(collapse = ',')) %>% unlist() %>% str_c(collapse = ';')
    }
  }),
  .after = 'FEATURE_PROTEIN_mapped'
)
dfs4c4 %>% rio::export('//172.16.13.136/share/members/jiangwenhao/TPHP/20220908/TPHP_carcinoma_upregulated_overlapped_drugbank/TableS4C-ANOVAsignificantHits_addcolumn_gene_protein_organ.xlsx')



```



```{r}
prot_bre_end <- intersect(tbl_ls$uterus$prot, tbl_ls$`mammary gland`$prot) # endometrial/breast overlapp


```



```{r}
# save.image()
# load('.RData')

```



